---
title: "MS1220_NHANES_HCHSSOL_AnalysisCode"
author: "Jeanette Varela, email:jvarela@hsph.harvard.edu"
output: html_document
---

The code provided in this document is used for both the NHANES and HCHS/SOL dataset. Here we have included the NHANES example. To use this code on the HCHS/SOL data, you can import the data and change the variable names. 

# Data Wrangling

```{r,echo=FALSE, warning=FALSE, message=FALSE}
#Libraries needed
require(dplyr)
require(stringr)
require(moments)
require(forcats)
require(naniar)
require(plyr)
require(survey)
require(psych)
require(jtools)
require(ggplot2)

```

## Nutrition variables

This code is taking the average of nutrition intake values for day 1 and day 2. Outliers will be removed and the log version will be used for analysis.

```{r,echo=FALSE, warning=FALSE, message=FALSE}
# Import data
Full_Data <- read.csv("~/OneDrive - Harvard University/CVD-Nutrient Research/Data/NHANES Data/Merged_NHANES_data.csv")

# Data wrangling
#combine SFA to Medium chain saturated fatty acids (MCSFA) and long chain saturated fatty acids (LCSFA) 
Full_Data$MCSFA_1 <- rowSums(Full_Data[, c("DR1TS060", "DR1TS080", "DR1TS100", "DR1TS120")])
Full_Data$LCSFA_1 <- rowSums(Full_Data[, c("DR1TS140", "DR1TS160", "DR1TS180")])

Full_Data$MCSFA_2 <- rowSums(Full_Data[, c("DR2TS060", "DR2TS080", "DR2TS100", "DR2TS120")])
Full_Data$LCSFA_2 <- rowSums(Full_Data[, c("DR2TS140", "DR2TS160", "DR2TS180")])

#drop SFA variables
delSFA <- c("DR1TS060", "DR1TS080", "DR1TS100", "DR1TS120","DR1TS140", "DR1TS160", "DR1TS180","DR2TS060", "DR2TS080", "DR2TS100", "DR2TS120","DR2TS140", "DR2TS160", "DR2TS180")
Full_Data <- Full_Data[,!(names(Full_Data) %in% delSFA)]

#keep only dietary recall = 1 (this indicates reliable data)
ReliRecall_Data  <- subset(Full_Data, DR1DRSTZ == 1 | DR2DRSTZ == 1)

# variables that will be kept 
keepvars <- c("SEQN", "RIAGENDR","RIDAGEYR", "RIDRETH1", "SDMVPSU", "SDMVSTRA", "WTINT2YR", "WTDRD1", "WTDR2D", "DMDEDUC2", "DMDYRSUS", "DMDBORN2", "USBORNLEN_C", "DMDHHSIZ", "INDHHIN2", "DMDMARTL","INDFMPIR", "OCQ380", "OCD150", "ALQ120Q", "ALQ101", "SMQ020", "SMQ040", "LBXTC", "LBDLDL", "LBDHDD", "RXDDRGID1", "RXDDRGID2", "RXDDRGID3", "RXDDRGID10", "RXDDRGID4", "RXDDRGID5", "RXDDRGID6", "RXDDRGID7", "RXDDRGID8", "RXDDRGID9", "RXDDRGID11", "RXDDRGID12", "RXDDRGID13", "RXDDRGID14", "RXDDRGID15", "RXDDRGID16", "RXDDRGID17", "RXDDRGID18", "DIQ280", "DIQ010", "PHAFSTHR", "LBXGLU", "LBXGLT", "DIQ050","DIQ070", "DID070", "BMXBMI", "BMXWAIST","RIAGENDR","RIDAGEYR","BPXSY1", "BPXSY2", "BPXSY3","BPXSY4", "BPXDI1", "BPXDI2", "BPXDI3","BPXDI4", "BPQ050A", "MCQ160B", "MCQ160C", "MCQ160D", "MCQ160E", "MCQ160F", "DR1TKCAL", "DR1TPROT", "DR1TCARB", "DR1TSUGR", "DR1TFIBE", "DR1TTFAT", "DR1TMFAT", "DR1TPFAT", "DR1TCHOL", "DR1TATOC", "DR1TRET",  "DR1TVARA",
 "DR1TACAR", "DR1TBCAR", "DR1TCRYP", "DR1TLYCO", "DR1TLZ",   "DR1TVB1", 
 "DR1TVB2",  "DR1TNIAC", "DR1TVB6",  "DR1TFOLA", "DR1TVB12", "DR1TVC",  
 "DR1TVD",   "DR1TVK",   "DR1TCALC", "DR1TPHOS", "DR1TMAGN", "DR1TIRON",
 "DR1TZINC", "DR1TCOPP", "DR1TSODI", "DR1TPOTA", "DR1TSELE", "DR1TCAFF",
 "DR1TS040", "MCSFA_1",  "LCSFA_1", "DR2TKCAL", "DR2TPROT",
 "DR2TCARB", "DR2TSUGR", "DR2TFIBE", "DR2TTFAT", "DR2TMFAT", "DR2TPFAT",
 "DR2TCHOL", "DR2TATOC", "DR2TRET",  "DR2TVARA", "DR2TACAR", "DR2TBCAR",
 "DR2TCRYP", "DR2TLYCO", "DR2TLZ",   "DR2TVB1",  "DR2TVB2",  "DR2TNIAC",
 "DR2TVB6",  "DR2TFOLA", "DR2TVB12", "DR2TVC",   "DR2TVD",   "DR2TVK", "DR2TCALC", "DR2TPHOS", "DR2TMAGN", "DR2TIRON", "DR2TZINC", "DR2TCOPP",
 "DR2TSODI", "DR2TPOTA", "DR2TSELE", "DR2TCAFF", "DR2TS040",
 "MCSFA_2",  "LCSFA_2")

#keep variables in the data set
ReliRecall_Data <- ReliRecall_Data[,(names(ReliRecall_Data) %in% keepvars)]
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=18, fig.height=10}
## Nutrient information
#need all nutrient variables, to compute the 2-day average

nutr <- c("DR1TKCAL", "DR1TPROT", "DR1TCARB", "DR1TSUGR", "DR1TFIBE", "DR1TTFAT",
 "DR1TMFAT", "DR1TPFAT", "DR1TCHOL", "DR1TATOC", "DR1TRET",  "DR1TVARA",
 "DR1TACAR", "DR1TBCAR", "DR1TCRYP", "DR1TLYCO", "DR1TLZ",   "DR1TVB1", 
 "DR1TVB2",  "DR1TNIAC", "DR1TVB6",  "DR1TFOLA", "DR1TVB12", "DR1TVC",  
 "DR1TVD",   "DR1TVK",   "DR1TCALC", "DR1TPHOS", "DR1TMAGN", "DR1TIRON",
 "DR1TZINC", "DR1TCOPP", "DR1TSODI", "DR1TPOTA", "DR1TSELE", "DR1TCAFF",
 "DR1TS040", "MCSFA_1",  "LCSFA_1", "DR2TKCAL", "DR2TPROT",
 "DR2TCARB", "DR2TSUGR", "DR2TFIBE", "DR2TTFAT", "DR2TMFAT", "DR2TPFAT",
 "DR2TCHOL", "DR2TATOC", "DR2TRET",  "DR2TVARA", "DR2TACAR", "DR2TBCAR",
 "DR2TCRYP", "DR2TLYCO", "DR2TLZ",   "DR2TVB1",  "DR2TVB2",  "DR2TNIAC",
 "DR2TVB6",  "DR2TFOLA", "DR2TVB12", "DR2TVC", "DR2TVD",   "DR2TVK", "DR2TCALC", "DR2TPHOS", "DR2TMAGN", "DR2TIRON", "DR2TZINC", "DR2TCOPP",
 "DR2TSODI", "DR2TPOTA", "DR2TSELE", "DR2TCAFF", "DR2TS040",
 "MCSFA_2",  "LCSFA_2")


nutr_data_D1.D2 <- ReliRecall_Data[,names(ReliRecall_Data) %in% nutr]

#rename the new variables MCSFA and LCSFA to keep it consistent with the dataset
nutr_data_D1.D2 <- nutr_data_D1.D2 %>% dplyr::rename(
      D1_MCSFA = MCSFA_1 ,
      D1_LCSFA = LCSFA_1 ,
      D2_MCSFA = MCSFA_2 ,
      D2_LCSFA = LCSFA_2 
    )

# make an indicator if it was 1 day or 2 days of nutrient data provided (this will help with the weights)

# everyone has first day of data so only checking the second day
NAs_day <- data.frame(rowSums(is.na(nutr_data_D1.D2[,c(38:74, 77,78)])))
useweights <- NULL
ind_missday2 <- ifelse(NAs_day != 0, 1, 0)

for(i in 1:nrow(nutr_data_D1.D2)){
  useweights <- ifelse(ind_missday2 == 1, ReliRecall_Data$WTDRD1, ReliRecall_Data$WTDR2D)
}

# empty vector for the 2-day mean for nutrient data
mean2days <- NULL

# functions computes the average for the two days
for (i in unique(colnames(nutr_data_D1.D2[,c(1:37, 75, 76)]))){
  for (j in unique(colnames(nutr_data_D1.D2[,c(38:74, 77,78)]))){
 if(stringr::str_sub(i, 4, 15) == stringr::str_sub(j, 4, 15)){
      mean2days <- rbind(mean2days, rowMeans(subset(nutr_data_D1.D2, select = c(i, j)), na.rm = TRUE))
    }
  }
}

#save the means in a dataframe
mean2days <- t(mean2days) #transpose matrix
df_mean <- data.frame(mean2days) # create a dataframe of values 
names(df_mean)[1:length(df_mean)] <- colnames(nutr_data_D1.D2[,c(1:37, 75,76)]) #add column names to the dataframe

dff <- df_mean

##### get rid of outliers in the data#####
#####delete those < 0.5th percentile and > 99.5th percentile in energy intake distribution

ENGYquantiles <- quantile( dff$DR1TKCAL, c(0.005, 0.995), na.rm =TRUE)

dff$DR1TKCAL <- ifelse(dff$DR1TKCAL < ENGYquantiles[1] ,0, dff$DR1TKCAL)
dff$DR1TKCAL <- ifelse(dff$DR1TKCAL > ENGYquantiles[2] ,0, dff$DR1TKCAL)

dff$SEQN <- ReliRecall_Data$SEQN

# add the weights to the dataframe
dff$WTDIETC <- useweights

# remove people that have extreme energy
dff <- subset(dff, dff$DR1TKCAL != 0)

#log version of nutrients, except energy
logdf <- dff
logdf[,2:39] <- data.matrix(log(dff[,2:39] + 0.1)) 
log_mean_nutr <- as.data.frame(logdf)
remove(logdf)

# regular units of nutrients
nutr_nonlog_mean <- dff

## need to integrate averaged nutrient intake variables back in to the data set but will delete the day 1 and day 2 nutrient measurements and add the correct weights
EditedNutr_Data <- ReliRecall_Data[,!names(ReliRecall_Data) %in% nutr]

EditedNutr_Data <- merge(EditedNutr_Data, log_mean_nutr, by="SEQN" )

EditedNutr_Data <-  EditedNutr_Data[complete.cases(EditedNutr_Data[,68:106]), ]# everyone had complete data
```

## Subsetting

Subsetting the data to the population we are interested in studying-- Mexican Americans and Other Hispanics ages 20-74. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#keep those ages 20-74
Data_Adults <- subset(EditedNutr_Data, RIDAGEYR > 19 & RIDAGEYR < 75 )

#keep Mexican American and Other Hispanic only
Data_Adults_MAOH <- subset(Data_Adults, RIDRETH1 < 3)

Final_Data <- Data_Adults_MAOH
```


## Recode Variables

```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width = "100%"}

#recode education variable to three levels - less than HS, HS diploma and more than HS 

Final_Data$EDU_C <- Final_Data$DMDEDUC2 

Final_Data <- Final_Data  %>% naniar::replace_with_na(replace = list(EDU_C = c(7,9)))

Final_Data <- Final_Data %>%
  mutate(EDU_C = as.factor(EDU_C) %>% 
           forcats::fct_recode("1" = "1", # less than HS
                               "1" = "2",
                               
                               "2" = "3", #HS
  
                               "3" = "4", #More than HS
                               "3" = "5"))

Final_Data$EDU_C = relevel(Final_Data$EDU_C, ref=3)


#recode length in the US into binary for less than or equal to 10 years and more than 10 years
Final_Data$USYR_C <- Final_Data$DMDYRSUS

Final_Data <- Final_Data %>% 
              replace_with_na(replace = list(USYR_C = c(77, 99)))

Final_Data$USYR_C <- ifelse( Final_Data$USYR_C < 4, 1, 2) #1 is less than 10 years, 2 is more than 10 years

Final_Data$USYR_C <- relevel(as.factor(Final_Data$USYR_C), ref = 2)

#recode born in US into binary variable for 1 born in the US and not born in the US
Final_Data$USBorn_C <- Final_Data$DMDBORN2

Final_Data <- Final_Data %>% 
              replace_with_na(replace = list(USBorn_C = c(7, 9)))

Final_Data$USBorn_C <- ifelse( Final_Data$USBorn_C == 1, 1, 2) # 1 is born in US, 2 is born in Mexico, Other Spanish Speaking Country or Other Non-Spanish Speaking Country 

Final_Data$USBorn_C = relevel(as.factor(Final_Data$USBorn_C), ref = 2)

# recode combination of US born and length in US
Final_Data <- Final_Data %>%
  mutate(USBORNLEN_C = factor(case_when((USBorn_C == 1) ~ "BORNUS",
                                (USYR_C == 1 ) ~ "Less10",
                                (USYR_C == 2 ) ~ "More10"),
                      levels = c("BORNUS", "Less10", "More10")))

Final_Data$USBORNLEN_C = relevel(as.factor(Final_Data$USBORNLEN_C), ref="BORNUS")


#recode number of people in household into binary variable with less than 2 or >= 2 
Final_Data$PPLHS_C <- Final_Data$DMDHHSIZ

Final_Data$PPLHS_C <- ifelse(Final_Data$PPLHS_C ==1, 1, 2) #1 indicates 1 or less and 2 indicates 2 or more

#recode household income
Final_Data$HSINC <- Final_Data$INDHHIN2

Final_Data <- Final_Data  %>% 
              replace_with_na(replace = list(HSINC = c(77, 99)))

Final_Data <- Final_Data %>%
  mutate(HSINC_C = factor(
    case_when((HSINC < 6) ~ "Less than $25,000",
             (HSINC == 6 | HSINC == 7 | HSINC == 8 | HSINC == 9 | HSINC == 10) ~ "$25,000-$75,000",
             (HSINC > 13) ~ "More than $75,000"),
              levels = c("Less than $25,000", "$25,000-$75,000", "More than $75,000")))



#recode martial status to binary married and not married
Final_Data$MS_C <- Final_Data$DMDMARTL

Final_Data <- Final_Data  %>% 
  replace_with_na(replace = list(MS_C = c(77, 99)))

Final_Data$MS_C <- ifelse((Final_Data$MS_C == 1 | Final_Data$MS_C == 6) , 1, 0) # 1 indicates married or living with partner

         
#recode poverty index ratio into a 3 categorical variable
Final_Data$INDPIR_C <- Final_Data$INDFMPIR

Final_Data$INDPIR_C <- 
  cut(as.numeric(Final_Data$INDPIR_C), 
      breaks = c(0, 1.3, 1.85, 6), 
      label =c("1", "2", "3") , right = FALSE)

Final_Data$INDPIR_C <- relevel(Final_Data$INDPIR_C, ref = 3)

#recode employment into 3 categorial variable with retired, employed and unemployed
Final_Data <- Final_Data  %>% 
              replace_with_na(replace = list(OCD150 = 7))

Final_Data <- Final_Data %>%
  mutate(EMP_C = factor(
    case_when((OCQ380 == 3 & OCD150 == 4) ~ "Retired",
              (OCD150 == 1 | OCD150 == 2) ~ "Employed",
              (OCD150 == 3) ~ "Unemployed",
              (OCQ380 != 3 & OCD150 == 4) ~ "Unemployed"),
              levels = c("Retired", "Employed", "Unemployed")))


Final_Data$EMP_C <- relevel(Final_Data$EMP_C, ref = "Retired")

#recode alcohol into 3 categories - never, current and former
Final_Data <- Final_Data  %>% 
  replace_with_na(replace = list(ALQ120Q = 999))

Final_Data <- Final_Data %>%
  mutate(ALC_C = factor(case_when((ALQ101 == 2) ~ "Never",
                                 (ALQ101 == 1 & ALQ120Q == 0) ~ "Former",
                                 (ALQ101 == 1 & ALQ120Q > 0) ~ "Current"),
                      levels = c("Never", "Former", "Current")))

Final_Data$ALC_C = relevel(Final_Data$ALC_C, ref = "Never")


#recode SMOKING into 3 categories - never, current and former

Final_Data <- Final_Data %>% 
  replace_with_na(replace = list(SMQ020 = 9))

Final_Data <- Final_Data %>%
  mutate(SM3_C = factor(
    case_when((SMQ020 == 2 ) ~ "Never",
              (SMQ020 == 1 & SMQ040  == 3) ~ "Former",
              (SMQ020 == 1 & SMQ040  < 3) ~ "Current"),
               levels = c("Never", "Former", "Current")))

#change into binary to match SOL/HCHS
Final_Data$SM_C <- ifelse(Final_Data$SM3_C == "Never" | Final_Data$SM3_C == "Former", 0, 1)

#Final_Data$SM_C <- relevel(Final_Data$SM_C, ref = "Never")


```

```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width = "100%"}
#Recode CVD risk factors

### Get cholesterol risk cutoff values from [National Heart, Lung, and Blood Institute](https://medlineplus.gov/cholesterollevelswhatyouneedtoknow.html)

Final_Data <- Final_Data %>%
  mutate(LBXTC_C = factor(case_when((LBXTC < 200) ~ "Normal",
                                    (LBXTC >= 200) ~ "High"),
                          levels = c("Normal", "High"))) %>%
  mutate(LBDLDL_C = factor(case_when((LBDLDL < 100) ~ "Normal",
                                     (LBDLDL >= 100) ~ "High"),
                           levels = c("Normal", "High"))) %>%
  mutate(LBDHDD_C = factor(case_when((LBDHDD > 45) ~ "Normal",
                                     (LBDHDD <= 45) ~ "Low"),
                           levels = c("Normal", "Low")))

############
#need to find chol meds for patient to better identify them

antilipid_names <- c("ATORVASTATIN", "FLUVASTATIN",
  "LOVASTATIN", "PRAVASTATIN", 
  "SIMVASTATIN", "FENOFIBRATE", 
  "GEMFIBROZIL", "CLOFIBRATE", 
  "COLESEVELAM", "COLESTIPOL", 
  "CHOLESTYRAMINE", "NIACIN", 
  "EZETIMIBE", "ORLISTAT")  

antilipid <- c("d04105", "d03183", "d00280","d04787", "d00348", "d00746", "d05348",  "d07110", "d07805", "d04286", "d00245", "d00196", "d04695", "d00744", "d00193", "d04787", "d07110", "d04824","d05348", "d04429")  

Final_Data$ind_CHOLMED <-
  seq(9, 9, length.out = length(Final_Data$SEQN))

Final_Data$ind_CHOLMED <-
  ifelse(Final_Data$RXDDRGID1 %in% antilipid |
      Final_Data$RXDDRGID2 %in% antilipid |
      Final_Data$RXDDRGID3 %in% antilipid |
      Final_Data$RXDDRGID10 %in% antilipid |
      Final_Data$RXDDRGID4 %in% antilipid |
      Final_Data$RXDDRGID5 %in% antilipid |
      Final_Data$RXDDRGID6 %in% antilipid |
      Final_Data$RXDDRGID7 %in% antilipid |
      Final_Data$RXDDRGID8 %in% antilipid |
      Final_Data$RXDDRGID9 %in% antilipid |
      Final_Data$RXDDRGID11 %in% antilipid |
      Final_Data$RXDDRGID12 %in% antilipid |
      Final_Data$RXDDRGID13 %in% antilipid |
      Final_Data$RXDDRGID14 %in% antilipid |
      Final_Data$RXDDRGID15 %in% antilipid |
      Final_Data$RXDDRGID16 %in% antilipid |
      Final_Data$RXDDRGID17 %in% antilipid |
      Final_Data$RXDDRGID18 %in% antilipid, 1, 0)



#########

Final_Data <- Final_Data %>% 
  mutate(CHOL_C = factor(
    ifelse((LBXTC >= 240 | LBDLDL >= 160 | LBDHDD < 40 & LBDHDD >= 0 | ind_CHOLMED ==1),
    "High Cholesterol", "No Cholesterol"), 
    levels = c("High Cholesterol", "No Cholesterol")))

## Fasting Glucose/Diabetes
### Get glucose risk categories from the [CDC](https://www.mayoclinic.org/diseases-conditions/diabetes/diagnosis-treatment/drc-20371451)

# correct the NAs in the diabetes variables that I need 
Final_Data <- Final_Data  %>%
  replace_with_na(replace = list(DIQ280 = c(777, 999)))

Final_Data <- Final_Data  %>% 
  replace_with_na(replace = list(DIQ010 = c(7, 9)))

Final_Data$DIQ010 <- ifelse(Final_Data$DIQ010 == 2 | Final_Data$DIQ010 == 3, 2, 1)


Final_Data$DIAB_test <- ifelse(is.na(Final_Data$DIQ280) & 
                                 is.na(Final_Data$PHAFSTHR) & 
                                 is.na(Final_Data$LBXGLU) & 
                                 is.na(Final_Data$LBXGLT) & 
                                 is.na(Final_Data$DIQ050) & 
                                 is.na(Final_Data$DID070) & 
                                 is.na(Final_Data$DIQ070) & 
                                 is.na(Final_Data$DIQ010), 99, 0)

Final_Data$DIAB_test1 <- ifelse((Final_Data$PHAFSTHR > 8 &
                                 Final_Data$LBXGLU >= 126) | 
                                (Final_Data$PHAFSTHR <= 8 & 
                                 Final_Data$LBXGLU >= 200) | 
                                 Final_Data$LBXGLT >= 200 | 
                                 Final_Data$DIQ280 >= 6.5, 1, 0) ##lab results

Final_Data$DIAB_test2 <- ifelse(Final_Data$DIQ010 == 1, 1, 0) # told diabetes by doctor, only 3 NAs

Final_Data$DIAB_test3 <- ifelse(Final_Data$DIQ050 == 1 | 
                                Final_Data$DID070 == 1 | 
                                Final_Data$DIQ070 == 1, 1, 0) #diabetes med use

Final_Data$DIAB_test4 <-
  rowSums(cbind(as.numeric(Final_Data$DIAB_test),
                as.numeric(Final_Data$DIAB_test1), 
                as.numeric(Final_Data$DIAB_test2),
                as.numeric(Final_Data$DIAB_test3)), na.rm = TRUE)

Final_Data$DIAB_C <- ifelse(Final_Data$DIAB_test4 == 0, 0, 1)

## BMI
### Get adult risk categories [from the CDC](https://www.cdc.gov/healthyweight/assessing/bmi/adult_bmi/index.html) 
Final_Data$BMI_C <- ifelse(Final_Data$BMXBMI >= 30, 1, 0)# non-obese -> 0, obese -> 1

#waist circumference 
Final_Data$WCM_C <- ifelse(Final_Data$BMXWAIST > 88 & 
                           Final_Data$RIAGENDR == 2, 1,
                           ifelse(Final_Data$BMXWAIST > 102 & 
                                  Final_Data$RIAGENDR == 1, 1, 0))


########### use this variable for in the analysis
# new obesity variable with BMI for 20-44 and Waist circum 45-75
Final_Data$OBS_C <- ifelse(Final_Data$RIDAGEYR < 45 & 
                           Final_Data$BMI_C == 1 , 1, 
                           ifelse(Final_Data$RIDAGEYR > 44 & 
                                  Final_Data$WCM_C == 1, 1, 0))

## Waist to hip ratio (not using)
#Final_Data$WHR <- Final_Data$BMXWAIST/Final_Data$BMX

#Final_Data$WHR_C <- ifelse(Final_Data$WHR > 88 &
#                           Final_Data$RIAGENDR == 2, 1, 
#                            ifelse(Final_Data$WHR > 102 & 
#                                   Final_Data$RIAGENDR == 1, 1, 0))


## Blood Pressure
Final_Data$AVG_SYS <- rowMeans(subset(Final_Data, select = 
                             c(BPXSY1, BPXSY2, BPXSY3, BPXSY4)), na.rm = TRUE)

Final_Data$AVG_DIA <- rowMeans(subset(Final_Data, select = 
                             c(BPXDI1, BPXDI2, BPXDI3, BPXDI4)), na.rm = TRUE)

Final_Data <- Final_Data %>%
  mutate(BP_C2 = factor(
    case_when((AVG_SYS < 120 & AVG_DIA < 80) ~ "Normal/Elevated",                                        (AVG_SYS < 130 & AVG_SYS >= 120 & AVG_DIA < 80) ~ "Normal/Elevated",
             ((AVG_SYS >= 130 & AVG_SYS < 180) | (AVG_DIA >= 80 & AVG_DIA < 120)) ~ "High/Crisis",(AVG_SYS >= 180 | AVG_DIA >= 120) ~ "High/Crisis"),
              levels = c("Normal/Elevated", "High/Crisis")))

# use this version--- it will be coded the same as hyper_treatment in HCHS/SOL 
# HCHS/SOL defined hypertension as med use and/or >= 140/90 mm Hg

Final_Data$HYPER_C <- ifelse((Final_Data$AVG_SYS >= 140 | 
                              Final_Data$AVG_DIA >= 90 | 
                              Final_Data$BPQ050A == 1 ), 1, 0)

Final_Data$HYPER_C <- ifelse(is.na(Final_Data$HYPER_C) == TRUE, 0, Final_Data$HYPER_C)

#BPQ050A - Now taking prescribed medicine for HBP

# Create Heart Disease variable like CVD_Frame in HCHS/SOL 
# CHD (coronary death, myocardial infarction, coronary insufficiency, and angina), cerebrovascular events (including ischemic stroke, hemorrhagic stroke, and transient ischemic attack), peripheral artery disease (intermittent claudication), and heart failure].  

# MHEA3. Has a doctor ever said that you have angina? ---- yes MCQ160D
# MHEA4. Has a doctor ever said that you have had a heart attack? ----- yes MCQ160E
# MHEA5. Has a doctor ever said that you had heart failure? ---- yes MCQ160B
# MHEA6. Has a doctor ever said that you had rheumatic heart disease?  ---- maybe MCQ160C
# MHEA7. Has a doctor ever told you that you had atrial fibrillation? -- not in NHANES
# MHEA8. Has a doctor ever said that you had some other kind of heart problem? -- no only for 99-04
# MHEA9. Have you had a balloon angioplasty, a stent, or bypass surgery to the arteries in your heart to improve the blood flow to your heart? ---- no???
# MHEA10. Has a doctor ever said that you had a stroke? ----- yes MCQ160F
# MHEA11. Has a doctor ever said that you had a mini-stroke of TIA (transient ischemic attack)? --SPQ070d
# MHEA12. Have you had a balloon angioplasty or surgery to the arteries of your neck to prevent or correct a stroke? --- OHQ146 but only from 99-04
# MHEA13. Has a doctor ever said that you have an aortic aneurysm, an AAA, or ballooning of your aorta? ----- no
# MHEA14. Has a doctor ever said that you have peripheral arterial disease (problems with circulation, blocked arteries to the legs)?----no
# ABIGRP2_C4. 4-level grouped Ankle Brachial Index – v2
# CLAUDICATION_INT. Intermittent Claudication - either leg
# MI_ECG. MI defined by ECG results

Final_Data <- Final_Data  %>% naniar::replace_with_na(replace = list(MCQ160B = c(7,9)))
Final_Data <- Final_Data  %>% naniar::replace_with_na(replace = list(MCQ160C = c(7,9)))
Final_Data <- Final_Data  %>% naniar::replace_with_na(replace = list(MCQ160D = c(7,9)))
Final_Data <- Final_Data  %>% naniar::replace_with_na(replace = list(MCQ160E = c(7,9)))
Final_Data <- Final_Data  %>% naniar::replace_with_na(replace = list(MCQ160F = c(7,9)))

Final_Data <- Final_Data %>%
  mutate(HD_C = factor(
    case_when((MCQ160B == 1 |MCQ160C==1 |MCQ160D==1|MCQ160E==1|MCQ160F==1 ) ~ "Heart Disease", (MCQ160B != 1 |MCQ160C != 1 |MCQ160D!=1|MCQ160E!=1|MCQ160F!=1) ~ "No Heart Disease"), levels = c("Heart Disease", "No Heart Disease")))
```



```{r}

#get cvd risk score 

# use package from github
#Calculate AHA/ACC 2013 ASCVD risk scores
library(devtools)

#install_github("vcastro/CVRisk")
library(CVrisk)
library(dplyr)
library(survey)
library(jtools)

Final_Data$Sex <- factor(Final_Data$RIAGENDR, levels=c(1,2), labels=c("male", "female"))

raceth <- c("Mexican","Other Hispanic", "NH White", "NH Black", "Other/Mixed")
Final_Data$Race <- factor(Final_Data$RIDRETH1, levels=c(1:5), labels=raceth)


Final_Data$bp_med <- Final_Data$BPQ050A
Final_Data$bp_med[is.na(Final_Data$bp_med)] <- 0


#Final_Data$Sex <- as.numeric(Final_Data$Sex)
Final_Data$Race <- as.character(Final_Data$Race)
Final_Data$SM_C <- as.numeric(Final_Data$SM_C)
Final_Data$CHOL_C <- as.numeric(Final_Data$CHOL_C)
Final_Data$DIAB_C <- as.numeric(Final_Data$DIAB_C)
Final_Data$BMI_C <- as.numeric(Final_Data$BMI_C)
Final_Data$bp_med <- as.numeric(Final_Data$bp_med)


# find how many NAs
#sapply(Final_Data[, c("RIDAGEYR", "Race","Sex","BMXBMI","AVG_SYS", "LBDHDD", "LBXTC", "BPQ050A",  "SM_C", "DIAB_C", "bp_med")], function(x) sum(is.na(x)))

#Framingham 2008 ASCVD risk score = ascvd_10y_frs
#Framingham 2008 ASCVD risk score simple (no lab) = ascvd_10y_frs_simple

# this has a lot of missing because of the bp_med variable having a lot of missing data
Final_Data_1 <- compute_CVrisk(Final_Data, scores="ascvd_10y_frs", age = "RIDAGEYR", race = "Race", gender ="Sex", bmi = "BMXBMI", sbp = "AVG_SYS", hdl = "LBDHDD", totchol = "LBXTC", bp_med = "bp_med", smoker = "SM_C", diabetes = "DIAB_C")


```

## Subset of dataset (no CVD)

```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width = "100%"}
# excluding patients with cvd  

nocvdpast <- subset(Final_Data_1, HD_C == "No Heart Disease") 

```


```{r}
nocvdpast$RIAGENDR <- as.factor(nocvdpast$RIAGENDR)
nocvdpast$RIDRETH1 <- as.factor(nocvdpast$RIDRETH1)
nocvdpast$EDU_C <- as.factor(nocvdpast$EDU_C)
nocvdpast$HSINC_C <- as.factor(nocvdpast$HSINC_C)
nocvdpast$PPLHS_C <- as.factor(nocvdpast$PPLHS_C)
nocvdpast$MS_C <- as.factor(nocvdpast$MS_C)
nocvdpast$ALC_C <- as.factor(nocvdpast$ALC_C)
nocvdpast$SM_C <- as.factor(nocvdpast$SM_C)
nocvdpast$CHOL_C <- as.factor(nocvdpast$CHOL_C)
nocvdpast$DIAB_C <- as.factor(nocvdpast$DIAB_C)
nocvdpast$BMI_C <- as.factor(nocvdpast$BMI_C)
```


# Factor Analysis

```{r, echo=FALSE}
# Use the data labeled nocvdpast

# variables needed in the dataset
vars <- c("SEQN", "RIAGENDR","RIDAGEYR", "RIDRETH1", "SDMVPSU", "SDMVSTRA", "WTDIETC",  "EDU_C", "USBORNLEN_C", "PPLHS_C", "HSINC_C", "MS_C", "EMP_C", "ALC_C", "SM_C", "CHOL_C", "DIAB_C", "OBS_C", "HYPER_C","ascvd_10y_frs", "DR1TKCAL", "DR1TPROT", "DR1TCARB", "DR1TSUGR", "DR1TFIBE", "DR1TTFAT", "DR1TMFAT", "DR1TPFAT", "DR1TCHOL", "DR1TATOC", "DR1TRET",  "DR1TVARA", "DR1TACAR", "DR1TBCAR", "DR1TCRYP", "DR1TLYCO", "DR1TLZ",   "DR1TVB1", "DR1TVB2",  "DR1TNIAC", "DR1TVB6",  "DR1TFOLA", "DR1TVB12", "DR1TVC",  
 "DR1TVD", "DR1TVK", "DR1TCALC", "DR1TPHOS", "DR1TMAGN", "DR1TIRON",
 "DR1TZINC", "DR1TCOPP", "DR1TSODI", "DR1TPOTA", "DR1TSELE", "DR1TCAFF",
 "DR1TS040", "D1_MCSFA", "D1_LCSFA")

AnalysisData <- nocvdpast[ ,(names(nocvdpast) %in% vars)]

numnames <- c("Total Protein (gm)", 
 "Total Carbohydrate (gm)",  "Total Sugars (gm)",       
 "Total Dietary Fiber (gm)", "Total Fat (gm)",          
 "Total MUFA (gm)",          "Total PUFA (gm)",         
 "Cholesterol (mg)",         "Vitamin E (mg)"  ,        
 "Retinol (mcg)"    ,        "Vitamin A (mcg)"  ,       
 "Alpha-Carotene (mcg)",     "Beta-Carotene (mcg)",     
 "Beta-Cryptoxanthin (mcg)", "Lycopene (mcg)"      ,    
 "Lutein+Zeaxanthin (mcg)" , "Vitamin B1 (mg)"      ,   
 "Vitamin B2 (mg)"          ,"Niacin (mg)",             
 "Vitamin B6 (mg)"          ,"Total Folate (mcg)",      
 "Vitamin B12 (mcg)"        ,"Vitamin C (mg)"  ,        
 "Vitamin D (mcg)"          ,"Vitamin K (mcg)" ,        
 "Calcium (mg)"             ,"Phosphorus (mg)",         
 "Magnesium (mg)"           ,"Iron (mg)" ,             
 "Zinc (mg)"                ,"Copper (mg)" ,            
 "Sodium (mg)"              ,"Potassium (mg)",          
 "Selenium (mcg)"           ,"Caffeine (mg)",           
 "SFA4 (gm)"                ,"MCSFA (gm)"  ,            
 "LCSFA (gm)" )

```



```{r}
# adjust for three cycles of NHANES data
AnalysisData$WTDIETC <- AnalysisData$WTDIETC* 1/3

# setting up the survey design
survey_design <- survey::svydesign(id = ~SDMVPSU,
                            strata = ~SDMVSTRA,
                            weights = ~WTDIETC,
                            nest = TRUE,
                            data = AnalysisData)

```


```{r}
#5 factors

svyfact5 <-
  survey::svyfactanal(
    ~ DR1TPROT + DR1TCARB + DR1TSUGR + DR1TFIBE + DR1TTFAT + DR1TMFAT + DR1TPFAT +
      DR1TCHOL + DR1TATOC + DR1TRET + DR1TVARA + DR1TACAR + DR1TBCAR + DR1TCRYP +
      DR1TLYCO + DR1TLZ + DR1TVB1 + DR1TVB2 + DR1TNIAC + DR1TVB6 + DR1TFOLA +
      DR1TVB12 + DR1TVC + DR1TVD + DR1TVK + DR1TCALC + DR1TPHOS + DR1TMAGN + DR1TIRON +
      DR1TZINC + DR1TCOPP + DR1TSODI + DR1TPOTA + DR1TSELE + DR1TCAFF + DR1TS040 +
      D1_MCSFA + D1_LCSFA,
    design = survey_design,
    factors = 5,
    lower = 0.1
  )


comm_tot5 <-
  matrix(apply(svyfact5$loadings[, 1:2], 1, function(x)
    sum(x ^ 2)),
    dim(AnalysisData[, 9:46])[2],
    1) # save the commonality values

svyfact_load5 <-
  as.data.frame(cbind(svyfact5$loadings[, 1:5], comm_tot5)) #save into dataframe
svyfact_load5$vars <- numnames # add variable name column
svyfact_load5 <-
  svyfact_load5[c(7, 1, 2, 3, 4, 5, 6)] # reorder the columns in dataframe


fact5_scores <-
  psych::factor.scores(AnalysisData[9:46], svyfact5$loadings, method = "Bartlett") # save the loading values and rotate

```



## Five Factor loadings

```{r}
# nutrients in factor 1
five_f1 <- svyfact_load5$vars[which(svyfact_load5$Factor1 >= 0.6)]

# nutrients in factor 2
five_f2 <- svyfact_load5$vars[which(svyfact_load5$Factor2 >= 0.6)]

# nutrients in factor 3
five_f3 <- svyfact_load5$vars[which(svyfact_load5$Factor3 >= 0.6)]

# nutrients in factor 4
five_f4 <- svyfact_load5$vars[which(svyfact_load5$Factor4 >= 0.6)]

# nutrients in factor 5
five_f5 <- svyfact_load5$vars[which(svyfact_load5$Factor5 >= 0.6)]
```



### Heat map

```{r}


#creating a heatmap for five factors
labs <- c( "Factor 1", "Factor 2", "Factor 3", "Factor 4", "Factor 5")

library(RColorBrewer)
coul <- colorRampPalette(brewer.pal(8, "Blues"))(25)

heatmap(as.matrix(cbind(svyfact_load5[,2:6])), Colv = NA, Rowv = NA, scale="column",col= coul, labRow= numnames, labCol = "",add.expr = text(x = seq_along(labs), y = 0, srt = 0, labels = labs, xpd = TRUE), main = "NHANES (5 Factors)")


# export to make a combined heat map

#write.csv(svyfact_load5, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Factor Analysis/NHANES_5svyfactor_10.23.23.csv")

```





## Combine data and factor scores -- 5 factors

```{r}
# factor scores are added to the full dataframe
Full_Data_5svyF <- cbind(AnalysisData, fact5_scores$scores)

#save a csv file of the data if needed
#write.csv(Full_Data_5svyF, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Factor Analysis/Full_Data_5svyF_10.23.23.csv")
```



### Cronbach's alpha

```{r}
# Survey weighted cronbach alpha - Nutrients ##

#Create survey object for HCHS nutrient data
nhanes.svynut <- svydesign(id = ~SDMVPSU,                      
                      strata  = ~SDMVSTRA,                    
                      weights = ~WTDIETC,                 
                      nest    = TRUE,                      
                      data    = AnalysisData)

#scale nutrients within survey design 
scale.svynut <-
  gscale(
    nhanes.svynut,
    vars = c(
      "DR1TPROT", "DR1TCARB", "DR1TSUGR", "DR1TFIBE",
      "DR1TTFAT", "DR1TMFAT", "DR1TPFAT", "DR1TCHOL",
      "DR1TATOC", "DR1TRET" , "DR1TVARA", "DR1TACAR",
      "DR1TBCAR", "DR1TCRYP", "DR1TLYCO", "DR1TLZ",
      "DR1TVB1" , "DR1TVB2" , "DR1TNIAC", "DR1TVB6",
      "DR1TFOLA", "DR1TVB12", "DR1TVC"  , "DR1TVD",
      "DR1TVK"  , "DR1TCALC", "DR1TPHOS", "DR1TMAGN",
      "DR1TIRON", "DR1TZINC", "DR1TCOPP", "DR1TSODI",
      "DR1TPOTA", "DR1TSELE", "DR1TCAFF", "DR1TS040",
      "D1_MCSFA", "D1_LCSFA"))


#Common factors from phif41_MSFA.xlsx - 16APR2021 all nutrients included
allfac <-
  svycralpha(
    ~ DR1TPROT + DR1TCARB + DR1TSUGR + DR1TFIBE + DR1TTFAT + DR1TMFAT + DR1TPFAT +
      DR1TCHOL + DR1TATOC + DR1TRET + DR1TVARA + DR1TACAR +
      DR1TBCAR + DR1TCRYP + DR1TLYCO + DR1TLZ + DR1TVB1 + DR1TVB2 +
      DR1TNIAC + DR1TVB6 + DR1TFOLA + DR1TVB12 + DR1TVC + DR1TVD +
      DR1TVK + DR1TCALC + DR1TPHOS + DR1TMAGN + DR1TIRON + DR1TZINC +
      DR1TCOPP + DR1TSODI + DR1TPOTA + DR1TSELE + DR1TCAFF + DR1TS040 + D1_MCSFA +
      D1_LCSFA,
    scale.svynut) 

#leave one out alpha
alpha.itemout <- function(a) {
  len <- dim(a)[1]
  a_out <- rep(NA, len)
  for (i in 1:len) {
    redmat <- a[-i, -i]
    cbar <- mean(redmat[lower.tri(redmat)])
    vbar <- mean(diag(redmat))
    a_outnum <- (len * cbar)
    a_outdem <- vbar + (len - 1) * cbar
    a_out[i] <- a_outnum / a_outdem
  }
  print(a_out)
}
  
  
#Step 1 - create covariance matrix
var.out <-
  svyvar(
    ~ DR1TPROT + DR1TCARB + DR1TSUGR + DR1TFIBE + DR1TTFAT + DR1TMFAT + DR1TPFAT +
      DR1TCHOL + DR1TATOC + DR1TRET + DR1TVARA + DR1TACAR + DR1TBCAR +
      DR1TCRYP + DR1TLYCO + DR1TLZ + DR1TVB1 + DR1TVB2 +
      DR1TNIAC + DR1TVB6 + DR1TFOLA + DR1TVB12 + DR1TVC + DR1TVD +
      DR1TVK + DR1TCALC + DR1TPHOS + DR1TMAGN + DR1TIRON + DR1TZINC +
      DR1TCOPP + DR1TSODI + DR1TPOTA + DR1TSELE + DR1TCAFF + DR1TS040 + D1_MCSFA +
      D1_LCSFA,
    scale.svynut)

covmat <- as.matrix(var.out)

fullvec <- alpha.itemout(covmat)

overall_alpha <- as.data.frame(cbind(nutr = c("Overall", numnames), val = round(c(allfac, fullvec), 3)), row.names = F)
```


```{r}
################################################
# 5 factors, compute Cronbach alpha for each factor
################################################

#rownames(svyfact_load5[which(svyfact_load5$Factor1 >= 0.4),])
#only include nutrients with at least 0.4 in phi_41 - factor 1
bay.5fac1 <- svycralpha(
  ~ DR1TPROT + DR1TCHOL + DR1TVB1 + DR1TVB2 + DR1TNIAC +
    DR1TVB6 + DR1TFOLA + DR1TVB12 + DR1TVD + DR1TPHOS +
    DR1TMAGN + DR1TIRON + DR1TZINC +
    DR1TCOPP + DR1TSODI + DR1TPOTA + DR1TSELE,
  scale.svynut)

fac1_5.var <- svyvar(
  ~ DR1TPROT + DR1TCHOL + DR1TVB1 + DR1TVB2 + DR1TNIAC +
    DR1TVB6 + DR1TFOLA + DR1TVB12 + DR1TVD + DR1TPHOS +
    DR1TMAGN + DR1TIRON + DR1TZINC +
    DR1TCOPP + DR1TSODI + DR1TPOTA + DR1TSELE,
  scale.svynut)

fac1_5.cov <- as.matrix(fac1_5.var)
fac1_5.alpha <- alpha.itemout(fac1_5.cov)

fact1_5alpha <-
  as.data.frame(cbind(nutr = c("Overall", numnames[which(svyfact_load5$Factor1 >= 0.4)]), val = round(c(bay.5fac1, fac1_5.alpha), 3)), row.names = F)

#rownames(svyfact_load5[which(svyfact_load5$Factor2 >= 0.4),])
#only include nutrients with at least 0.4 in phi_41 - factor 2
fac2_5 <-
  svycralpha(
    ~ DR1TPROT + DR1TTFAT + DR1TMFAT + DR1TPFAT + DR1TCHOL + DR1TATOC + DR1TPHOS +
      DR1TZINC + DR1TSODI + DR1TSELE + D1_LCSFA,
    scale.svynut)


fac2_5.var <-
  as.matrix(
    svyvar(
      ~ DR1TPROT + DR1TTFAT + DR1TMFAT + DR1TPFAT + DR1TCHOL + DR1TATOC + DR1TPHOS +
        DR1TZINC + DR1TSODI + DR1TSELE + D1_LCSFA,
      scale.svynut))

fac2_5.alpha <- alpha.itemout(fac2_5.var)

fact2_5alpha <-
  as.data.frame(cbind(nutr = c("Overall", numnames[which(svyfact_load5$Factor2 >= 0.4)]), val = round(c(fac2_5, fac2_5.alpha), 3)), row.names = F)


#rownames(svyfact_load5[which(svyfact_load5$Factor3 >= 0.4),])
#only include nutrients with at least 0.4 in phi_41 - factor 3
fac3_5 <-
  svycralpha(
    ~ DR1TCARB + DR1TSUGR + DR1TFIBE + DR1TATOC + DR1TVB1 + DR1TFOLA + DR1TCALC +
      DR1TPHOS + DR1TMAGN + DR1TIRON + DR1TCOPP + DR1TPOTA,
    scale.svynut)

fac3_5.var <-
  as.matrix(
    svyvar(
      ~ DR1TCARB + DR1TSUGR + DR1TFIBE + DR1TATOC + DR1TVB1 + DR1TFOLA + DR1TCALC +
        DR1TPHOS + DR1TMAGN + DR1TIRON + DR1TCOPP + DR1TPOTA,
      scale.svynut))

fac3_5.alpha <- alpha.itemout(fac3_5.var)

fact3_5alpha <-
  as.data.frame(cbind(nutr = c("Overall", numnames[which(svyfact_load5$Factor3 >= 0.4)]), val = round(c(fac3_5, fac3_5.alpha), 3)), row.names = F)


#rownames(svyfact_load5[which(svyfact_load5$Factor4 >= 0.4),])
#only include nutrients with at least 0.4 in phi_41 - factor 4
fac4_5 <-
  svycralpha(
    ~ DR1TRET + DR1TVARA + DR1TVB2 + DR1TVB12 + DR1TVD + DR1TCALC + DR1TS040 +
      D1_MCSFA + D1_LCSFA,
    scale.svynut)

fac4_5.var <-
  as.matrix(
    svyvar(
      ~ DR1TRET + DR1TVARA + DR1TVB2 + DR1TVB12 + DR1TVD + DR1TCALC + DR1TS040 +
        D1_MCSFA + D1_LCSFA,
      scale.svynut))

fac4_5.alpha <- alpha.itemout(fac4_5.var)

fact4_5alpha <-
  as.data.frame(cbind(nutr = c("Overall", numnames[which(svyfact_load5$Factor4 >= 0.4)]), val = round(c(fac4_5, fac4_5.alpha), 3)), row.names = F)


#rownames(svyfact_load5[which(svyfact_load5$Factor5 >= 0.4),])
#only include nutrients with at least 0.4 in phi_41 - factor 5
fac5_5 <-
  svycralpha(
    ~ DR1TATOC + DR1TVARA + DR1TACAR + DR1TBCAR + DR1TCRYP + DR1TLZ + DR1TVC +
      DR1TVK + DR1TPOTA,
    scale.svynut)

fac5_5.var <-
  as.matrix(
    svyvar(
      ~ DR1TATOC + DR1TVARA + DR1TACAR + DR1TBCAR + DR1TCRYP + DR1TLZ + DR1TVC +
        DR1TVK + DR1TPOTA,
      scale.svynut))

fac5_5.alpha <- alpha.itemout(fac5_5.var)

fact5_5alpha <-
  as.data.frame(cbind(nutr = c("Overall", numnames[which(svyfact_load5$Factor5 >= 0.4)]), val = round(c(fac5_5, fac5_5.alpha), 3)), row.names = F)


### Table of Cronbach's alpha

alpha_5 <- list(overall_alpha, fact1_5alpha, fact2_5alpha, fact3_5alpha, fact4_5alpha, fact5_5alpha) %>% purrr::reduce(dplyr::left_join, by = "nutr")

colnames(alpha_5) <- c("Nutrient","Overall", "Factor 1", "Factor 2", "Factor 3", "Factor 4", "Factor 5")

#write.csv(alpha_5, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/NHANES/NHANES_5svyfactor_Cronbachs_10.23.23.csv")
```

# Descriptive Analysis

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# use data labeled Full_Data_5svyF

vars <-
  c(
    "SEQN",
    "RIAGENDR",
    "RIDAGEYR",
    "RIDRETH1",
    "SDMVPSU",
    "SDMVSTRA",
    "WTDIETC",
    "EDU_C",
    "USBORNLEN_C",
    "PPLHS_C",
    "HSINC_C",
    "MS_C",
    "EMP_C",
    "ALC_C",
    "SM_C",
    "CHOL_C",
    "DIAB_C",
    "OBS_C",
    "ascvd_10y_frs",
    "HYPER_C",
    "DR1TKCAL",
    "DR1TPROT",
    "DR1TCARB",
    "DR1TSUGR",
    "DR1TFIBE",
    "DR1TTFAT",
    "DR1TMFAT",
    "DR1TPFAT",
    "DR1TCHOL",
    "DR1TATOC",
    "DR1TRET",
    "DR1TVARA",
    "DR1TACAR",
    "DR1TBCAR",
    "DR1TCRYP",
    "DR1TLYCO",
    "DR1TLZ",
    "DR1TVB1",
    "DR1TVB2",
    "DR1TNIAC",
    "DR1TVB6",
    "DR1TFOLA",
    "DR1TVB12",
    "DR1TVC",
    "DR1TVD",
    "DR1TVK",
    "DR1TCALC",
    "DR1TPHOS",
    "DR1TMAGN",
    "DR1TIRON",
    "DR1TZINC",
    "DR1TCOPP",
    "DR1TSODI",
    "DR1TPOTA",
    "DR1TSELE",
    "DR1TCAFF",
    "DR1TS040",
    "D1_MCSFA",
    "D1_LCSFA",
    "Factor1",
    "Factor2",
    "Factor3",
    "Factor4",
    "Factor5"
  )

variables <- c("ID", "Sex", "Age (yrs)", "Ethnicity/Race", "PSU", "Strata", "Weights", "Education", "Born Length", "Ppl in Household", "Household Income", "Marital Status", "Employment", "Alcohol status", "Smoking status", "Cholesterol", "Diabetes", "Obesity", "Hypertension","Fram_score", "Energy (kcal)", "Total Protein (gm)", "Total Carbohydrate (gm)", "Total Sugars (gm)", "Total Dietary Fiber (gm)", "Total Fat (gm)", "Total MUFA (gm)", "Total PUFA (gm)", "Cholesterol (mg)", "Vitamin E (mg)", "Retinol (mcg)", "Vitamin A (mcg)", "Alpha-Carotene (mcg)", "Beta-Carotene (mcg)", "Beta-Cryptoxanthin (mcg)", "Lycopene (mcg)", "Lutein+Zeaxanthin (mcg)", "Vitamin B1 (mg)", "Vitamin B2 (mg)", "Niacin (mg)", "Vitamin B6 (mg)", "Total Folate (mcg)", "Vitamin B12 (mcg)", "Vitamin C (mg)", "Vitamin D (mcg)", "Vitamin K (mcg)", "Calcium (mg)", "Phosphorus (mg)", "Magnesium (mg)", "Iron (mg)", "Zinc (mg)", "Copper (mg)", "Sodium (mg)", "Potassium (mg)", "Selenium (mcg)", "Caffeine (mg)", "SFA4 (butyric acid) (gm)", "MCSFA (gm)", "LCSFA (gm)", "Factor1", "Factor2", "Factor3", "Factor4", "Factor5")

TableData <- Full_Data_5svyF[,(names(Full_Data_5svyF) %in% vars)]
AnalysisData <- Full_Data_5svyF[,(names(Full_Data_5svyF) %in% vars)]


# table of demographics and cvd risk factors
# want to make table pretty

TableData$RIAGENDR <- ifelse(TableData$RIAGENDR == 2, "Female", "Male")
TableData$RIDRETH1 <- ifelse(TableData$RIDRETH1 == 2, "Other Hispanic", "Mexican")
TableData$EDU_C <- ifelse(TableData$EDU_C == 1, "Less than HS", ifelse(TableData$EDU_C ==2, "HS Diploma", "More than HS"))
TableData$USBORNLEN_C <- ifelse(TableData$USBORNLEN_C == "BORNUS", "Born in the US", ifelse(TableData$USBORNLEN_C =="Less10", "Less than 11 years", "More than 10 years"))
TableData$PPLHS_C <- ifelse(TableData$PPLHS_C == 2, "2 or more people", "Less than 2 people")
TableData$MS_C <- ifelse(TableData$MS_C == 1, "Married", "Unmarried")
TableData$SM_C <- ifelse(TableData$SM_C == 1, "Smoker", "Non-smoker")
TableData$CHOL_C <- ifelse(TableData$CHOL_C == 1, "Yes", "No")
TableData$DIAB_C <- ifelse(TableData$DIAB_C == 1, "Yes", "No")
TableData$OBS_C <- ifelse(TableData$OBS_C == 1, "Yes", "No")
TableData$HYPER_C <- ifelse(TableData$HYPER_C == 1, "Yes", "No")


TableData$RIAGENDR <- as.factor(TableData$RIAGENDR)
TableData$RIDRETH1 <- as.factor(TableData$RIDRETH1)
TableData$EDU_C <- as.factor(TableData$EDU_C)
TableData$USBORNLEN_C <- as.factor(TableData$USBORNLEN_C)
TableData$PPLHS_C <- as.factor(TableData$PPLHS_C)
TableData$HSINC_C <- as.factor(TableData$HSINC_C)
TableData$MS_C <- as.factor(TableData$MS_C)
TableData$EMP_C <- as.factor(TableData$EMP_C)
TableData$ALC_C <- as.factor(TableData$ALC_C)
TableData$SM_C <- as.factor(TableData$SM_C)
TableData$CHOL_C <- as.factor(TableData$CHOL_C)
TableData$DIAB_C <- as.factor(TableData$DIAB_C)
TableData$OBS_C <- as.factor(TableData$OBS_C)
TableData$HYPER_C <- as.factor(TableData$HYPER_C)


```


```{r, echo=FALSE}
# setting up the survey design for the table 
survey_design <- survey::svydesign(id = ~SDMVPSU,
                            strata = ~SDMVSTRA,
                            weights = ~WTDIETC,
                            nest = TRUE,
                            data = TableData)

demovars <- c("RIAGENDR","RIDAGEYR","DR1TKCAL", "RIDRETH1","EDU_C", "USBORNLEN_C", "PPLHS_C", "HSINC_C", "MS_C", "EMP_C", "ALC_C", "SM_C", "CHOL_C", "DIAB_C", "OBS_C", "HYPER_C","ascvd_10y_frs" )

```


## Demographics Table (Age Adjusted)

```{r, echo=FALSE}
# using this example to derive age adjusted summaries
#https://github.com/mlaviolet/Survey-data-analysis/blob/master/age-adjusted-prevalence.R
# use svystandardize to age adjust and new package srvyr 

# https://seer.cancer.gov/stdpopulations/stdpop.singleages.html to get population numbesr
pop4 <- c(40376, 37648, 35557, 34892, 34099, 34211, 33282, 34350, 34506, 40872, 39990, 38102, 37744,38409, 40869, 42881, 43496, 44695, 42902, 47826, 46667, 44936, 44876, 44240, 44074, 42680,40339, 39585, 36815, 38640, 37209, 35043, 34757, 37542, 27692, 27497, 27868, 29475, 24045,24188, 22591, 21798, 21329, 20307, 20518, 20339, 18621, 18499, 17888, 18752, 18431, 17847,18021, 16743, 16214)

# setting up the adjusted survey design
age_adj <- svystandardize(
  survey_design,
  by = ~ RIDAGEYR,
  over = ~ 1,
  population = pop4
) 

##########################
#compute the mean using svymean with individual variable

scsloop <- function(vars, design){ 
svymean(as.formula(paste0( "~" ,vars)) ,design ,na=T)}

xmeans2 <- lapply(demovars, scsloop, age_adj)

tab1_adj <- plyr::ldply(xmeans2, data.frame)
tab1_adj [3,2] <- tab1_adj [3,3]
tab1_adj [4,2] <- tab1_adj [4,4]
tab1_adj [36,2] <- tab1_adj [36,5]

tab1_adj <- tab1_adj[1:36, 1:2]

RN_1 <- NULL
for(i in 1:17){
  RN_1 <- c(RN_1, names(xmeans2[[i]]))
}

rownames(tab1_adj) <- RN_1
blank <- rep("", 36)

#add a column with demo categories
rowname <- c("Female", "Male", "Age (yrs)", "Energy (kcal)", "Mexican", "Other Hispanic", "HS Diploma/GED", "< HS DiplomaGED", "> HS DiplomaGED", "Born in the US", "<= 10 years", "> 10 years", "2 or more people", "Less than 2 people", "$25,000-$75,000", "< $25,000", "> $75,000", "Married", "Unmarried", "Employed", "Retired", "Unemployed", "Current", "Former", "Never", "Non-smoker", "Smoker", "No", "Yes", "No", "Yes", "No", "Yes", "No", "Yes","Risk Score")

table1_adj <- cbind(blank, rowname, tab1_adj)

table1_adj[,3:4] <- round(table1_adj[,3:4],4)*100

#write.csv(table1_adj, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Descriptive Analysis/NHANES_DEMO_table_10.23.23.csv")

```



## Nutrition Table (Age,Sex, Energy Adjusted)


```{r, echo=FALSE}

###### using svyglm to get means of adjusted nutrient values
nutrvars <- vars[22:59]

geo_mean <- matrix(NA, length(nutrvars), 3)

for(i in 1:length(nutrvars)){
  mod1 <- svyglm(as.formula(paste0(nutrvars[i], "~" ,"RIDAGEYR + RIAGENDR + DR1TKCAL")), survey_design)
  mean_nutr <- as.data.frame(predict(mod1))
  geom <- as.data.frame(exp(mean_nutr$link))
  geo_mean[i,1] <- mean(geom$`exp(mean_nutr$link)`, na.rm=T)
  LC <- exp(mean_nutr$link - 1.96 * mean_nutr$SE)
  UC <- exp(mean_nutr$link + 1.96 * mean_nutr$SE)

  geo_mean[i,2] <- mean(LC)
  geo_mean[i,3] <- mean(UC)

}

geo_mean <- round(geo_mean, 1)


tab2_glm_adj_geo <- as.data.frame(paste0(geo_mean[,1], " (" ,geo_mean[,2], ", ", geo_mean[,3], ")"))


#write.csv(tab2_glm_adj_geo, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Descriptive Analysis/NHANES_NUTR_table_10.23.23.csv")
```



## Quintiles using survey quantile

```{r, echo=FALSE}

# Creating quintiles of the factor loadings
# factor 1
fac1 <- svyquantile( ~ Factor1, survey_design, c(.2, .4, .6, .8))
fac1.quints <- fac1$Factor1[1:4, 1]
AnalysisData$fac1_cat <-
  ifelse(AnalysisData$Factor1 <= fac1.quints[1], 1,
    ifelse(AnalysisData$Factor1 > fac1.quints[1] &
            AnalysisData$Factor1 < fac1.quints[2], 2,
      ifelse(AnalysisData$Factor1 >= fac1.quints[2] &
              AnalysisData$Factor1 < fac1.quints[3], 3,
        ifelse(AnalysisData$Factor1 >= fac1.quints[3] &
               AnalysisData$Factor1 < fac1.quints[4], 4,
          ifelse(AnalysisData$Factor1 >= fac1.quints[4], 5, NA)))))

AnalysisData$fac1_cat <-
  factor(AnalysisData$fac1_cat,
         c(1:5),
         labels = c("Q1", "Q2", "Q3", "Q4", "Q5"))
TableData$fac1_cat <- AnalysisData$fac1_cat

#factor 2
fac2 <- svyquantile( ~ Factor2, survey_design, c(.2, .4, .6, .8))
fac2.quints <- fac2$Factor2[1:4, 1]
AnalysisData$fac2_cat <-
  ifelse(AnalysisData$Factor2 <= fac2.quints[1], 1,
    ifelse(AnalysisData$Factor2 > fac2.quints[1] &
            AnalysisData$Factor2 < fac2.quints[2], 2,
      ifelse(AnalysisData$Factor2 >= fac2.quints[2] &
              AnalysisData$Factor2 < fac2.quints[3], 3,
        ifelse(AnalysisData$Factor2 >= fac2.quints[3] &
                AnalysisData$Factor2 < fac2.quints[4], 4,
          ifelse(AnalysisData$Factor2 >= fac2.quints[4], 5, NA)))))

AnalysisData$fac2_cat <-
  factor(AnalysisData$fac2_cat,
         c(1:5),
         labels = c("Q1", "Q2", "Q3", "Q4", "Q5"))
TableData$fac2_cat <- AnalysisData$fac2_cat


#factor 3
fac3 <- svyquantile( ~ Factor3, survey_design, c(.2, .4, .6, .8))
fac3.quints <- fac3$Factor3[1:4, 1]
AnalysisData$fac3_cat <-
  ifelse(AnalysisData$Factor3 <= fac3.quints[1], 1,
    ifelse(AnalysisData$Factor3 > fac3.quints[1] &
            AnalysisData$Factor3 < fac3.quints[2], 2,
      ifelse(AnalysisData$Factor3 >= fac3.quints[2] &
              AnalysisData$Factor3 < fac3.quints[3], 3,
        ifelse(AnalysisData$Factor3 >= fac3.quints[3] &
                AnalysisData$Factor3 < fac3.quints[4], 4,
          ifelse(AnalysisData$Factor3 >= fac3.quints[4], 5, NA)))))

AnalysisData$fac3_cat <-
  factor(AnalysisData$fac3_cat,
         c(1:5),
         labels = c("Q1", "Q2", "Q3", "Q4", "Q5"))

TableData$fac3_cat <- AnalysisData$fac3_cat


#factor 4
fac4 <- svyquantile( ~ Factor4, survey_design, c(.2, .4, .6, .8))
fac4.quints <- fac4$Factor4[1:4, 1]
AnalysisData$fac4_cat <-
  ifelse(AnalysisData$Factor4 <= fac4.quints[1], 1,
    ifelse(AnalysisData$Factor4 > fac4.quints[1] &
            AnalysisData$Factor4 < fac4.quints[2], 2,
      ifelse(AnalysisData$Factor4 >= fac4.quints[2] &
              AnalysisData$Factor4 < fac4.quints[3], 3,
        ifelse(AnalysisData$Factor4 >= fac4.quints[3] &
                AnalysisData$Factor4 < fac4.quints[4], 4,
          ifelse(AnalysisData$Factor4 >= fac4.quints[4], 5, NA)))))

AnalysisData$fac4_cat <-
  factor(AnalysisData$fac4_cat,
         c(1:5),
         labels = c("Q1", "Q2", "Q3", "Q4", "Q5"))

TableData$fac4_cat <- AnalysisData$fac4_cat


#factor 5
fac5 <- svyquantile( ~ Factor5, survey_design, c(.2, .4, .6, .8))
fac5.quints <- fac5$Factor5[1:4, 1]
AnalysisData$fac5_cat <-
  ifelse(AnalysisData$Factor5 <= fac5.quints[1], 1,
    ifelse(AnalysisData$Factor5 > fac5.quints[1] &
            AnalysisData$Factor5 < fac5.quints[2], 2,
      ifelse(AnalysisData$Factor5 >= fac5.quints[2] &
              AnalysisData$Factor5 < fac5.quints[3], 3,
        ifelse(AnalysisData$Factor5 >= fac5.quints[3] &
                AnalysisData$Factor5 < fac5.quints[4], 4,
          ifelse(AnalysisData$Factor5 >= fac5.quints[4], 5, NA)))))

AnalysisData$fac5_cat <-
  factor(AnalysisData$fac5_cat,
         c(1:5),
         labels = c("Q1", "Q2", "Q3", "Q4", "Q5"))

TableData$fac5_cat <- AnalysisData$fac5_cat
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# setting up the survey design
survey_design_Fortable <- survey::svydesign(
  id = ~ SDMVPSU,
  strata = ~ SDMVSTRA,
  weights = ~ WTDIETC,
  nest = TRUE,
  data = TableData)

#Demographic Variables from Table 1
  #FACTOR 1 - COMMON
tab_f1 <-
  as.data.frame(
    svyby(
      ~ RIAGENDR + RIDAGEYR + DR1TKCAL + RIDRETH1 + EDU_C + USBORNLEN_C + PPLHS_C +
        HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + OBS_C + HYPER_C +
        ascvd_10y_frs,
      ~ fac1_cat,
      survey_design_Fortable,
      svymean,
      na.rm = TRUE))

tab_f1 <- as.data.frame(t(tab_f1))

tab_f1$Q1 <-as.numeric(as.character(tab_f1$Q1))
tab_f1$Q2 <-as.numeric(as.character(tab_f1$Q2))
tab_f1$Q3 <-as.numeric(as.character(tab_f1$Q3))
tab_f1$Q4 <-as.numeric(as.character(tab_f1$Q4))
tab_f1$Q5 <-as.numeric(as.character(tab_f1$Q5))


#FACTOR 2 - COMMON
tab_f2 <-
  svyby(
    ~ RIAGENDR + RIDAGEYR + DR1TKCAL + RIDRETH1 + EDU_C + USBORNLEN_C + PPLHS_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + OBS_C + HYPER_C +
      ascvd_10y_frs,
    ~ fac2_cat,
    survey_design_Fortable,
    svymean,
    na.rm = TRUE)

tab_f2 <- as.data.frame(t(tab_f2))

tab_f2$Q1 <-as.numeric(as.character(tab_f2$Q1))
tab_f2$Q2 <-as.numeric(as.character(tab_f2$Q2))
tab_f2$Q3 <-as.numeric(as.character(tab_f2$Q3))
tab_f2$Q4 <-as.numeric(as.character(tab_f2$Q4))
tab_f2$Q5 <-as.numeric(as.character(tab_f2$Q5))

#FACTOR 3- COMMON
tab_f3 <-
  svyby(
    ~ RIAGENDR + RIDAGEYR + DR1TKCAL + RIDRETH1 + EDU_C + USBORNLEN_C + PPLHS_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + OBS_C + HYPER_C +
      ascvd_10y_frs,
    ~ fac3_cat,
    survey_design_Fortable,
    svymean,
    na.rm = TRUE)

tab_f3 <- as.data.frame(t(tab_f3))

tab_f3$Q1 <-as.numeric(as.character(tab_f3$Q1))
tab_f3$Q2 <-as.numeric(as.character(tab_f3$Q2))
tab_f3$Q3 <-as.numeric(as.character(tab_f3$Q3))
tab_f3$Q4 <-as.numeric(as.character(tab_f3$Q4))
tab_f3$Q5 <-as.numeric(as.character(tab_f3$Q5))

#FACTOR 4 - COMMON
tab_f4 <-
  svyby(
    ~ RIAGENDR + RIDAGEYR + DR1TKCAL + RIDRETH1 + EDU_C + USBORNLEN_C + PPLHS_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + OBS_C + HYPER_C +
      ascvd_10y_frs,
    ~ fac4_cat,
    survey_design_Fortable,
    svymean,
    na.rm = TRUE)

tab_f4 <- as.data.frame(t(tab_f4))

tab_f4$Q1 <-as.numeric(as.character(tab_f4$Q1))
tab_f4$Q2 <-as.numeric(as.character(tab_f4$Q2))
tab_f4$Q3 <-as.numeric(as.character(tab_f4$Q3))
tab_f4$Q4 <-as.numeric(as.character(tab_f4$Q4))
tab_f4$Q5 <-as.numeric(as.character(tab_f4$Q5))

#FACTOR 5 - COMMON
tab_f5 <-
  svyby(
    ~ RIAGENDR + RIDAGEYR + DR1TKCAL + RIDRETH1 + EDU_C + USBORNLEN_C + PPLHS_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + OBS_C + HYPER_C +
      ascvd_10y_frs,
    ~ fac5_cat,
    survey_design_Fortable,
    svymean,
    na.rm = TRUE)

tab_f5 <- as.data.frame(t(tab_f5))

tab_f5$Q1 <-as.numeric(as.character(tab_f5$Q1))
tab_f5$Q2 <-as.numeric(as.character(tab_f5$Q2))
tab_f5$Q3 <-as.numeric(as.character(tab_f5$Q3))
tab_f5$Q4 <-as.numeric(as.character(tab_f5$Q4))
tab_f5$Q5 <-as.numeric(as.character(tab_f5$Q5))
```

```{r, echo=FALSE}

table <- cbind.data.frame(
  Q1f1m = tab_f1[c(2:37), 1],
  Q1f1se = tab_f1[c(38:73), 1],
  Q5f1m = tab_f1[c(2:37), 5],
  Q5f1se = tab_f1[c(38:73), 5],
  Q1f2m = tab_f2[c(2:37), 1],
  Q1f2se = tab_f2[c(38:73), 1],
  Q5f2m = tab_f2[c(2:37), 5],
  Q5f2se = tab_f2[c(38:73), 5],
  Q1f3m = tab_f3[c(2:37), 1],
  Q1f3se = tab_f3[c(38:73), 1],
  Q5f3m = tab_f3[c(2:37), 5],
  Q5f3se = tab_f3[c(38:73), 5],
  Q1f4m = tab_f4[c(2:37), 1],
  Q1f4se = tab_f4[c(38:73), 1],
  Q5f4m = tab_f4[c(2:37), 5],
  Q5f4se = tab_f4[c(38:73), 5],
  Q1f5m = tab_f5[c(2:37), 1],
  Q1f5se = tab_f5[c(38:73), 1],
  Q5f5m = tab_f5[c(2:37), 5],
  Q5f5se = tab_f5[c(38:73), 5])

rownames(table) <- row.names(tab_f1)[2:37]
table <- round(table, 3) * 100

#write.csv(table, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Descriptive Analysis/NHANES_DEMO_quintiletable_10.23.23.csv")
```

# Regression Analysis

```{r}
# remove everything except data 
#rm(list=setdiff(ls(), "AnalysisData"))
# use AnalysisData
```

```{r, echo=FALSE}
# Need to re-level and categorize the  variables 

AnalysisData$USBORNLEN_C <-
  ifelse(AnalysisData$USBORNLEN_C == "BORNUS", 3,
    ifelse(AnalysisData$USBORNLEN_C == "Less10", 1, 2))

AnalysisData$HSINC_C <-
  ifelse(AnalysisData$HSINC_C == "Less than $25,000", 1,
    ifelse(AnalysisData$HSINC_C == "$25,000-$75,000", 2, 3))

AnalysisData$MS_C <-
  ifelse(AnalysisData$MS_C == 1, 2, 1) # 2 is unmarried

AnalysisData$EMP_C <-
  ifelse(AnalysisData$EMP_C == "Unemployed", 1,
    ifelse(AnalysisData$EMP_C == "Retired", 2, 3))

AnalysisData$ALC_C <-
  ifelse(AnalysisData$ALC_C == "Current", 1,
    ifelse(AnalysisData$ALC_C == "Former", 2, 3))

AnalysisData$SM_C <- ifelse(AnalysisData$SM_C == 1, 1, 2)
AnalysisData$OBS_C <- ifelse(AnalysisData$OBS_C == 1, 1, 2)
AnalysisData$HYPER_C <- ifelse(AnalysisData$HYPER_C == 1, 1, 2)
AnalysisData$DIAB_C <- ifelse(AnalysisData$DIAB_C == 1, 1, 2)

# make sure factor variables are specified using as.factor
AnalysisData$RIAGENDR <- as.factor(AnalysisData$RIAGENDR)
AnalysisData$RIDRETH1 <- as.factor(AnalysisData$RIDRETH1)
AnalysisData$EMP_C <- as.factor(AnalysisData$EMP_C)
AnalysisData$EDU_C <- as.factor(AnalysisData$EDU_C)
AnalysisData$MS_C <- as.factor(AnalysisData$MS_C)
AnalysisData$SM_C <- as.factor(AnalysisData$SM_C)
AnalysisData$ALC_C <- as.factor(AnalysisData$ALC_C)
AnalysisData$HSINC_C <- as.factor(AnalysisData$HSINC_C)
AnalysisData$USBORNLEN_C <- as.factor(AnalysisData$USBORNLEN_C)
AnalysisData$CHOL_C <- as.factor(AnalysisData$CHOL_C)
AnalysisData$OBS_C <- as.factor(AnalysisData$OBS_C)
AnalysisData$HYPER_C <- as.factor(AnalysisData$HYPER_C)
AnalysisData$DIAB_C <- as.factor(AnalysisData$DIAB_C)

# relevel to have the preferred reference level
AnalysisData$RIAGENDR <- relevel(AnalysisData$RIAGENDR, ref = 1) #female is  2
AnalysisData$RIDRETH1 <- relevel(AnalysisData$RIDRETH1, ref = 2) # Mexican is 1
AnalysisData$EMP_C <- relevel(AnalysisData$EMP_C, ref = 3)
#AnalysisData$EDU_C <- relevel(AnalysisData$EDU_C, ref = 3)
AnalysisData$MS_C <- relevel(AnalysisData$MS_C, ref = 2)
AnalysisData$SM_C <- relevel(AnalysisData$SM_C, ref = 2)
AnalysisData$ALC_C <- relevel(AnalysisData$ALC_C, ref = 3)
AnalysisData$HSINC_C <- relevel(AnalysisData$HSINC_C, ref = 3)
AnalysisData$USBORNLEN_C <- relevel(AnalysisData$USBORNLEN_C, ref =3)
AnalysisData$CHOL_C <- relevel(AnalysisData$CHOL_C, ref=2)
AnalysisData$OBS_C <- relevel(AnalysisData$OBS_C, ref = 2)
AnalysisData$HYPER_C <- relevel(AnalysisData$HYPER_C, ref = 2)
AnalysisData$DIAB_C <- relevel(AnalysisData$DIAB_C, ref = 2)

#setting up survey design but with a quintile 3 as the reference
survey_design_Q3 <- survey::svydesign(
  id = ~ SDMVPSU,
  strata = ~ SDMVSTRA,
  weights = ~ WTDIETC,
  nest = TRUE,
  data = AnalysisData)

```

```{r, echo=F, fig.width=10, fig.height=10, message=F, warning=F}

# relevel quintiles to set the reference to be level 3
survey_design_Q3[["variables"]][["fac1_cat"]] <- relevel(survey_design_Q3[["variables"]][["fac1_cat"]], ref = "Q3")

survey_design_Q3[["variables"]][["fac2_cat"]] <- relevel(survey_design_Q3[["variables"]][["fac2_cat"]], ref = "Q3")

survey_design_Q3[["variables"]][["fac3_cat"]] <- relevel(survey_design_Q3[["variables"]][["fac3_cat"]], ref = "Q3")

survey_design_Q3[["variables"]][["fac4_cat"]] <- relevel(survey_design_Q3[["variables"]][["fac4_cat"]], ref = "Q3")

survey_design_Q3[["variables"]][["fac5_cat"]] <- relevel(survey_design_Q3[["variables"]][["fac5_cat"]], ref = "Q3")
```

## NHANES Diabetes

```{r, echo=F, fig.width=10, fig.height=10, message=F, warning=F}
# Diabetes

#factor 1 - Q3 as ref
diabf1_refQ3 <-
  survey::svyglm(
    DIAB_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + OBS_C + HYPER_C + fac1_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

#####################################################################################

#factor 2 - Q3 as ref
diabf2_refQ3 <-
  survey::svyglm(
    DIAB_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + OBS_C + HYPER_C + fac2_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

#####################################################################################

#factor 3 - Q3 as ref
diabf3_refQ3 <-
  survey::svyglm(
    DIAB_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + OBS_C + HYPER_C + fac3_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

#####################################################################################
#factor 4 - Q3 as ref
diabf4_refQ3 <-
  survey::svyglm(
    DIAB_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + OBS_C + HYPER_C + fac4_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

#####################################################################################

#factor 5 - Q3 as ref
diabf5_refQ3 <-
  survey::svyglm(
    DIAB_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + OBS_C + HYPER_C + fac5_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

#####################################################################################
#all factors 
diabf15_Q3ref <-
  survey::svyglm(
    DIAB_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + OBS_C + HYPER_C + fac1_cat + fac2_cat + fac3_cat + fac4_cat + fac5_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

#####################################################################################
#https://stats.stackexchange.com/questions/97024/false-discovery-rate-of-multiple-regressions-models
#FDR - single models

diab_raw_pvals <- cbind(summary(diabf1_refQ3)$coefficients[,4],
              summary(diabf2_refQ3)$coefficients[,4],
              summary(diabf3_refQ3)$coefficients[,4],
              summary(diabf4_refQ3)$coefficients[,4],
              summary(diabf5_refQ3)$coefficients[,4])

diab_adj_pvals <- FDRestimation::p.fdr(p=diab_raw_pvals)
              
diab_adj_pvals_table <- matrix(diab_adj_pvals[["fdrs"]],ncol = 5)
rownames(diab_adj_pvals_table) <- rownames(diab_raw_pvals)


#####
#FDR - SINGLE AND ALL FACTOR MODELS


diab_raw_pvals_all <- as.matrix(c(summary(diabf1_refQ3)$coefficients[,4],
              summary(diabf2_refQ3)$coefficients[,4],
              summary(diabf3_refQ3)$coefficients[,4],
              summary(diabf4_refQ3)$coefficients[,4],
              summary(diabf5_refQ3)$coefficients[,4],
              summary(diabf15_Q3ref)$coefficients[,4]))

diab_adj_pvals_all <- FDRestimation::p.fdr(p=diab_raw_pvals_all)
              
diab_adj_pvals_table_all <- matrix(diab_adj_pvals_all[["fdrs"]], ncol=1)
rownames(diab_adj_pvals_table_all) <- rownames(diab_raw_pvals_all)

```


```{r, echo=F, fig.width=10, fig.height=10, message=F, warning=F}
### Plots

#factor 1 with Q3 as reference 
boxLabels_refQ3 = c(
  "Female",
  "Age",
  "Ethnicity(Mexican)",
  "Energy" ,
  "< HS/GED" ,
  "HS/GED",
  "US < 11 yrs",
  "US > 10 yrs",
  "Income < $25k",
  "Income $25-75k",
  "Unmarried",
  "Retired",
  "Unemployed",
  "Current Drinker",
  "Former Drinker",
  "Smoker",
  "Cholesterol",
  "Obesity",
  "Hypertension",
  "1st Quintile",
  "2nd Quintile",
  "4th Quintile",
  "5th Quintile"
)

# factor 1- save odds and CIs into dataframe to plot
df_f1_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(diabf1_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(diabf1_refQ3)[2:24]),
    boxCIHigh = exp(confint(diabf1_refQ3)[2:24, 2])
  )

pf1 <- ggplot(df_f1_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio") 


#factor 2 - save odds and CIs into dataframe to plot
df_f2_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(diabf2_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(diabf2_refQ3)[2:24]),
    boxCIHigh = exp(confint(diabf2_refQ3)[2:24, 2]))

pf2 <- ggplot(df_f2_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio") 


#factor 3- save odds and CIs into dataframe for plots
df_f3_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(diabf3_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(diabf3_refQ3)[2:24]),
    boxCIHigh = exp(confint(diabf3_refQ3)[2:24, 2])
  )

pf3 <- ggplot(df_f3_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")


#factor 4 - save odds and CIs into dataframe for plots

df_f4_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(diabf4_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(diabf4_refQ3)[2:24]),
    boxCIHigh = exp(confint(diabf4_refQ3)[2:24, 2])
  )

pf4 <- ggplot(df_f4_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio") 

#factor 5 - save odds and CIs into dataframe for plots

df_f5_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(diabf5_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(diabf5_refQ3)[2:24]),
    boxCIHigh = exp(confint(diabf5_refQ3)[2:24, 2])
  )

pf5 <- ggplot(df_f5_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio") 


#all factors with q3 as a reference
boxLabels15_Q3ref = c("Female", "Age", "Ethnicity(Mexican)","Energy" , "< HS/GED" ,"HS/GED", "US < 11 yrs", "US > 10 yrs", "Income < $25k",  "Income $25-75k", "Unmarried",  "Retired", "Unemployed", "Current Drinker", "Former Drinker", "Smoker", "Cholesterol", "Obesity", "Hypertension", "Meats-1st Quintle", "Meats-2nd Quintile", "Meats-4th Quintile", "Meats-5th Quintile", "Fats/Oils-1st Quintle", "Fats/Oils-2nd Quintile", "Fats/Oils-4th Quintile", "Fats/Oils-5th Quintile", "Grains/Legumes-1st Quintle", "Grains/Legumes-2nd Quintile", "Grains/Legumes-4th Quintile", "Grains/Legumes-5th Quintile","Dairy-1st Quintle", "Dairy-2nd Quintile", "Dairy-4th Quintile", "Dairy-5th Quintile", "Fruits/Veggies-1st Quintle", "Fruits/Veggies-2nd Quintile", "Fruits/Veggies-4th Quintile", "Fruits/Veggies-5th Quintile" )


df_f15_Q3ref <-
  data.frame(
    yAxis = length(boxLabels15_Q3ref):1,
    boxOdds = exp(diabf15_Q3ref$coefficients[2:40]),
    boxCILow = exp(confint(diabf15_Q3ref)[2:40, 1]),
    boxCIHigh = exp(confint(diabf15_Q3ref)[2:40, 2])
  )

pf15_Q3ref <-
  ggplot(df_f15_Q3ref, aes(x = boxOdds, y = boxLabels15_Q3ref)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")


ggpubr::ggarrange(pf1, pf2, pf3, pf4, pf5, pf15_Q3ref, labels = c("Meats", "Fats/Oils", "Grains/Legumes","Dairy", "Fruits/Veggies", "All Nutrient Patterns"), ncol = 2, nrow = 3)

```


```{r, echo=F,  fig.width=10, fig.height=10, message=F, warning=F}
#saving results into dataframe to export for comparison
NHANES_Diab_SingleFactors_refQ3 <- cbind(df_f1_refQ3, df_f2_refQ3, df_f3_refQ3, df_f4_refQ3, df_f5_refQ3)

#write.csv(df_f15_Q3ref, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Regression Analysis/NHANES_Diab_AllFactors_ORest_Q3ref_10.23.23.csv")

#write.csv(NHANES_Diab_SingleFactors_refQ3, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Regression Analysis/NHANES_Diab_SingleFactors_ORest_Q3ref_10.23.23.csv")
```

## NHANES Obesity

```{r, echo=F,  fig.width=10, fig.height=10, message=F, warning=F}
# Obesity

#factor 1 - Q3 ref
obsf1_refQ3 <-
  survey::svyglm(
    OBS_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + HYPER_C + fac1_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

###########################################################################################

#factor 2 - Q3 ref
obsf2_refQ3 <-
  survey::svyglm(
    OBS_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + HYPER_C + fac2_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

###########################################################################################

#factor 3 - Q3 ref
obsf3_refQ3 <-
  survey::svyglm(
    OBS_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + HYPER_C + fac3_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

###############################################################################################

#factor 4 - Q3 ref
obsf4_refQ3 <-
  survey::svyglm(
    OBS_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + HYPER_C + fac4_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

###############################################################################################

#factor 5 - Q3 ref
obsf5_refQ3 <-
  survey::svyglm(
    OBS_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + HYPER_C + fac5_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

###############################################################################################

#all factors with q3 as reference
obsf15_Q3ref <-
  survey::svyglm(
    OBS_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + HYPER_C + fac1_cat +fac1_cat +fac2_cat +fac3_cat +fac4_cat +fac5_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

#####################################################################################
#FDR - single models

obs_raw_pvals <- cbind(summary(obsf1_refQ3)$coefficients[,4],
              summary(obsf2_refQ3)$coefficients[,4],
              summary(obsf3_refQ3)$coefficients[,4],
              summary(obsf4_refQ3)$coefficients[,4],
              summary(obsf5_refQ3)$coefficients[,4])

obs_adj_pvals <- FDRestimation::p.fdr(p=obs_raw_pvals)
              
obs_adj_pvals_table <- matrix(obs_adj_pvals[["fdrs"]],ncol = 5)
rownames(obs_adj_pvals_table) <- rownames(obs_raw_pvals)

#####
#FDR - SINGLE AND ALL FACTOR MODELS

obs_raw_pvals_all <- as.matrix(c(summary(obsf1_refQ3)$coefficients[,4],
              summary(obsf2_refQ3)$coefficients[,4],
              summary(obsf3_refQ3)$coefficients[,4],
              summary(obsf4_refQ3)$coefficients[,4],
              summary(obsf5_refQ3)$coefficients[,4],
              summary(obsf15_Q3ref)$coefficients[,4]
              ))

obs_adj_pvals_all <- FDRestimation::p.fdr(p=obs_raw_pvals_all)
              
obs_adj_pvals_table_all <- matrix(obs_adj_pvals_all[["fdrs"]], ncol=1)
rownames(obs_adj_pvals_table_all) <- rownames(obs_raw_pvals_all)
```

```{r, echo=F,  fig.width=10, fig.height=10, message=F, warning=F}

#factor 1 with Q3 as reference
boxLabels_refQ3 = c("Female", "Age", "Ethnicity(Mexican)","Energy" , "< HS/GED" ,"HS/GED", "US < 11 yrs", "US > 10 yrs", "Income < $25k",  "Income $25-75k", "Unmarried",  "Retired", "Unemployed", "Current Drinker", "Former Drinker", "Smoker", "Cholesterol", "Diabetes", "Hypertension", "1st Quintile", "2nd Quintile","4th Quintile","5th Quintile" )

df_f1_obs_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(obsf1_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(obsf1_refQ3)[2:24]),
    boxCIHigh = exp(confint(obsf1_refQ3)[2:24, 2])
  )

pf1  <-
  ggplot(df_f1_obs_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")


#factor 2
df_f2_obs_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(obsf2_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(obsf2_refQ3)[2:24]),
    boxCIHigh = exp(confint(obsf2_refQ3)[2:24, 2])
  )

pf2 <-
  ggplot(df_f2_obs_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")

#factor 3
df_f3_obs_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(obsf3_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(obsf3_refQ3)[2:24]),
    boxCIHigh = exp(confint(obsf3_refQ3)[2:24, 2])
  )

pf3 <-
  ggplot(df_f3_obs_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")

#factor 4
df_f4_obs_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(obsf4_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(obsf4_refQ3)[2:24]),
    boxCIHigh = exp(confint(obsf4_refQ3)[2:24, 2])
  )

pf4 <-
  ggplot(df_f4_obs_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")

#factor 5
df_f5_obs_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(obsf5_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(obsf5_refQ3)[2:24]),
    boxCIHigh = exp(confint(obsf5_refQ3)[2:24, 2])
  )

pf5 <-
  ggplot(df_f5_obs_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio") 

#all factors with q3 as a reference
boxLabels15_Q3ref = c("Female", "Age", "Ethnicity(Mexican)","Energy" , "< HS/GED" ,"HS/GED", "US < 11 yrs", "US > 10 yrs", "Income < $25k",  "Income $25-75k", "Unmarried",  "Retired", "Unemployed", "Current Drinker", "Former Drinker", "Smoker", "Cholesterol", "Obesity", "Hypertension", "Meats-1st Quintle", "Meats-2nd Quintile", "Meats-4th Quintile", "Meats-5th Quintile", "Fats/Oils-1st Quintle", "Fats/Oils-2nd Quintile", "Fats/Oils-4th Quintile", "Fats/Oils-5th Quintile", "Grains/Legumes-1st Quintle", "Grains/Legumes-2nd Quintile", "Grains/Legumes-4th Quintile", "Grains/Legumes-5th Quintile", "Dairy-1st Quintle", "Dairy-2nd Quintile", "Dairy-4th Quintile", "Dairy-5th Quintile","Fruits/Veggies-1st Quintle", "Fruits/Veggies-2nd Quintile", "Fruits/Veggies-4th Quintile", "Fruits/Veggies-5th Quintile" )


df_f15_Q3ref <-
  data.frame(
    yAxis = length(boxLabels15_Q3ref):1,
    boxOdds = exp(obsf15_Q3ref$coefficients[2:40]),
    boxCILow = exp(confint(obsf15_Q3ref)[2:40, 1]),
    boxCIHigh = exp(confint(obsf15_Q3ref)[2:40, 2])
  )

pf15_Q3ref <-
  ggplot(df_f15_Q3ref, aes(x = boxOdds, y = boxLabels15_Q3ref)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio") 

# one plot for all nutrients
ggpubr::ggarrange(pf1, pf2, pf3, pf4, pf5, pf15_Q3ref, labels = c("Meats", "Fats/Oils", "Grains/Legumes","Dairy", "Fruits/Veggies", "All Nutrient Patterns"), ncol = 2, nrow = 3)
```


```{r, echo=F,  fig.width=10, fig.height=10, message=F, warning=F}
#saving results into dataframe to export for comparison
NHANES_Obesity_SingleFactors_refQ3 <- cbind(df_f1_obs_refQ3, df_f2_obs_refQ3, df_f3_obs_refQ3, df_f4_obs_refQ3, df_f5_obs_refQ3)

#write.csv(df_f15_Q3ref, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Regression Analysis/NHANES_Obesity_AllFactors_ORest_Q3ref_10.23.23.csv")

#write.csv(NHANES_Obesity_SingleFactors_refQ3, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Regression Analysis/NHANES_Obesity_SingleFactors_ORest_Q3ref_10.23.23.csv")
```

## NHANES Hypertension

```{r, echo=FALSE,  fig.width=10, fig.height=10, message=F, warning=F}
# Hypertension
#factor 1 - Q3 ref
hypf1_refQ3 <-
  survey::svyglm(
    HYPER_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + OBS_C + fac1_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

###########################################################################################

#factor 2 - Q3 ref
hypf2_refQ3 <-
  survey::svyglm(
    HYPER_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + OBS_C + fac2_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )
###########################################################################################

#factor 3 - Q3 ref
hypf3_refQ3 <-
  survey::svyglm(
    HYPER_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + OBS_C + fac3_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

###############################################################################################

#factor 4 - Q3 ref
hypf4_refQ3 <-
  survey::svyglm(
    HYPER_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + OBS_C + fac4_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

###############################################################################################

#factor 5 - Q3 ref
hypf5_refQ3 <-
  survey::svyglm(
    HYPER_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + OBS_C + fac5_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

###############################################################################################
#all factors with q3 as a reference
hypf15_Q3ref <-
  survey::svyglm(
    HYPER_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + CHOL_C + DIAB_C + OBS_C + fac1_cat + fac2_cat +fac3_cat +fac4_cat +fac5_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )


#####################################################################################
#FDR - single modles

hyp_raw_pvals <- cbind(summary(hypf1_refQ3)$coefficients[,4],
              summary(hypf2_refQ3)$coefficients[,4],
              summary(hypf3_refQ3)$coefficients[,4],
              summary(hypf4_refQ3)$coefficients[,4],
              summary(hypf5_refQ3)$coefficients[,4])

hyp_adj_pvals <- FDRestimation::p.fdr(p=hyp_raw_pvals)
              
hyp_adj_pvals_table <- matrix(hyp_adj_pvals[["fdrs"]],ncol = 5)
rownames(hyp_adj_pvals_table) <- rownames(hyp_raw_pvals)

#####
#FDR - SINGLE AND ALL FACTOR MODELS
hyp_raw_pvals_all <- as.matrix(c(summary(hypf1_refQ3)$coefficients[,4],
              summary(hypf2_refQ3)$coefficients[,4],
              summary(hypf3_refQ3)$coefficients[,4],
              summary(hypf4_refQ3)$coefficients[,4],
              summary(hypf5_refQ3)$coefficients[,4],
              summary(hypf15_Q3ref)$coefficients[,4]
              ))

hyp_adj_pvals_all <- FDRestimation::p.fdr(p=hyp_raw_pvals_all)
              
hyp_adj_pvals_table_all <- matrix(hyp_adj_pvals_all[["fdrs"]], ncol=1)
rownames(hyp_adj_pvals_table_all) <- rownames(hyp_raw_pvals_all)
```

```{r, echo=FALSE,  fig.width=10, fig.height=10, message=F, warning=F}
### Plots

#factor 1 Q3 ref
boxLabels_refQ3 = c("Female", "Age", "Ethnicity(Mexican)","Energy" , "< HS/GED" ,"HS/GED", "US < 11 yrs", "US > 10 yrs", "Income < $25k",  "Income $25-75k", "Unmarried",  "Retired", "Unemployed", "Current Drinker", "Former Drinker", "Smoker", "Cholesterol", "Diabetes", "Obesity", "1st Quintile", "2nd Quintile", "4th Quintile","5th Quintile" )

df_f1_hyp_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(hypf1_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(hypf1_refQ3)[2:24]),
    boxCIHigh = exp(confint(hypf1_refQ3)[2:24, 2])
  )

pf1  <-
  ggplot(df_f1_hyp_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")

#factor 2
df_f2_hyp_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(hypf2_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(hypf2_refQ3)[2:24]),
    boxCIHigh = exp(confint(hypf2_refQ3)[2:24, 2])
  )

pf2 <-
  ggplot(df_f2_hyp_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")

#factor 3
df_f3_hyp_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(hypf3_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(hypf3_refQ3)[2:24]),
    boxCIHigh = exp(confint(hypf3_refQ3)[2:24, 2])
  )

pf3 <-
  ggplot(df_f3_hyp_refQ3, aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")

#factor 4
df_f4_hyp_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(hypf4_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(hypf4_refQ3)[2:24]),
    boxCIHigh = exp(confint(hypf4_refQ3)[2:24, 2])
  )

pf4 <-
  ggplot(df_f4_hyp_refQ3 , aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")

#factor 5
df_f5_hyp_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(hypf5_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(hypf5_refQ3)[2:24]),
    boxCIHigh = exp(confint(hypf5_refQ3)[2:24, 2])
  )

pf5 <-
  ggplot(df_f5_hyp_refQ3 , aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio") 

#all factors with q3 as a reference
boxLabels15_Q3ref = c("Female", "Age", "Ethnicity(Mexican)","Energy" , "< HS/GED" ,"HS/GED", "US < 11 yrs", "US > 10 yrs", "Income < $25k",  "Income $25-75k", "Unmarried",  "Retired", "Unemployed", "Current Drinker", "Former Drinker", "Smoker", "Cholesterol", "Obesity", "Hypertension", "Meats-1st Quintle", "Meats-2nd Quintile", "Meats-4th Quintile", "Meats-5th Quintile", "Fats/Oils-1st Quintle", "Fats/Oils-2nd Quintile", "Fats/Oils-4th Quintile", "Fats/Oils-5th Quintile", "Grains/Legumes-1st Quintle", "Grains/Legumes-2nd Quintile", "Grains/Legumes-4th Quintile", "Grains/Legumes-5th Quintile","Dairy-1st Quintle", "Dairy-2nd Quintile", "Dairy-4th Quintile", "Dairy-5th Quintile", "Fruits/Veggies-1st Quintle", "Fruits/Veggies-2nd Quintile", "Fruits/Veggies-4th Quintile", "Fruits/Veggies-5th Quintile")


hypdf_f15_Q3ref <-
  data.frame(
    yAxis = length(boxLabels15_Q3ref):1,
    boxOdds = exp(hypf15_Q3ref$coefficients[2:40]),
    boxCILow = exp(confint(hypf15_Q3ref)[2:40, 1]),
    boxCIHigh = exp(confint(hypf15_Q3ref)[2:40, 2])
  )

pf15_Q3ref <-
  ggplot(hypdf_f15_Q3ref, aes(x = boxOdds, y = boxLabels15_Q3ref)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio") 

ggpubr::ggarrange(pf1, pf2, pf3, pf4, pf5, pf15_Q3ref, labels = c("Meats", "Fats/Oils", "Grains/Legumes", "Dairy", "Fruits/Veggies", "All Nutrient Patterns"), ncol = 2, nrow = 3)
```

```{r, echo=F,  fig.width=10, fig.height=10, message=F, warning=F}
#saving results into dataframe to export for comparison
NHANES_Hypert_SingleFactors_Q3 <- cbind(df_f1_hyp_refQ3, df_f2_hyp_refQ3, df_f3_hyp_refQ3, df_f4_hyp_refQ3, df_f5_hyp_refQ3)

#write.csv(hypdf_f15_Q3ref, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Regression Analysis/NHANES_Hypert_AllFactors_ORest_Q3ref_10.23.23.csv")

#write.csv(NHANES_Hypert_SingleFactors_Q3, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Regression Analysis/NHANES_Hypert_SingleFactors_ORest_Q3ref_10.23.23.csv")
```

\newpage

## NHANES Cholesterol

```{r, echo=FALSE,  fig.width=10, fig.height=10, message=F, warning=F}
# Cholesterol

#factor 1 - Q3 ref
cholf1_refQ3 <-
  survey::svyglm(
    CHOL_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + HYPER_C + DIAB_C + OBS_C + fac1_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )
###########################################################################################

#factor 2 - Q3 ref
cholf2_refQ3 <-
  survey::svyglm(
    CHOL_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + HYPER_C + DIAB_C + OBS_C + fac2_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

###############################################################################################

#factor 3 - Q3 ref
cholf3_refQ3 <-
  survey::svyglm(
    CHOL_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + HYPER_C + DIAB_C + OBS_C + fac3_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

###############################################################################################

#factor 4 - Q3 ref
cholf4_refQ3 <-
  survey::svyglm(
    CHOL_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + HYPER_C + DIAB_C + OBS_C + fac4_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )

###############################################################################################

#factor 5 - Q3 ref
cholf5_refQ3 <-
  survey::svyglm(
    CHOL_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + HYPER_C + DIAB_C + OBS_C + fac5_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )
###############################################################################################

#all factors with q3 as a reference
cholf15_Q3ref <-
  survey::svyglm(
    CHOL_C ~ RIAGENDR + RIDAGEYR + RIDRETH1 + DR1TKCAL + EDU_C + USBORNLEN_C +
      HSINC_C + MS_C + EMP_C + ALC_C + SM_C + HYPER_C + DIAB_C + OBS_C + fac1_cat + fac1_cat +fac2_cat +fac3_cat +fac4_cat +fac5_cat,
    design = survey_design_Q3,
    family = quasibinomial()
  )


#####################################################################################
#FDR - single modles

chol_raw_pvals <- cbind(summary(cholf1_refQ3)$coefficients[,4],
              summary(cholf2_refQ3)$coefficients[,4],
              summary(cholf3_refQ3)$coefficients[,4],
              summary(cholf4_refQ3)$coefficients[,4],
              summary(cholf5_refQ3)$coefficients[,4])

chol_adj_pvals <- FDRestimation::p.fdr(p=hyp_raw_pvals)
              
chol_adj_pvals_table <- matrix(chol_adj_pvals[["fdrs"]],ncol = 5)
rownames(chol_adj_pvals_table) <- rownames(chol_raw_pvals)

#####
#FDR - SINGLE AND ALL FACTOR MODELS
chol_raw_pvals_all <- as.matrix(c(summary(cholf1_refQ3)$coefficients[,4],
              summary(cholf2_refQ3)$coefficients[,4],
              summary(cholf3_refQ3)$coefficients[,4],
              summary(cholf4_refQ3)$coefficients[,4],
              summary(cholf5_refQ3)$coefficients[,4],
              summary(cholf15_Q3ref)$coefficients[,4]
              ))

chol_adj_pvals_all <- FDRestimation::p.fdr(p=chol_raw_pvals_all)
              
chol_adj_pvals_table_all <- matrix(chol_adj_pvals_all[["fdrs"]], ncol=1)
rownames(chol_adj_pvals_table_all) <- rownames(chol_raw_pvals_all)
```


```{r, echo=FALSE,  fig.width=10, fig.height=10, message=F, warning=F}
### Plots

#factor 1 with Q3 ref 
boxLabels_refQ3 = c("Female", "Age", "Ethnicity(Mexican)","Energy" , "< HS/GED" ,"HS/GED", "US < 11 yrs", "US > 10 yrs", "Income < $25k",  "Income $25-75k", "Unmarried",  "Retired", "Unemployed", "Current Drinker", "Former Drinker", "Smoker", "Hypertension", "Diabetes", "Obesity", "1st Quintile", "2nd Quintile", "4th Quintile","5th Quintile")

df_f1_chol_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(cholf1_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(cholf1_refQ3)[2:24]),
    boxCIHigh = exp(confint(cholf1_refQ3)[2:24, 2])
  )

pf1  <-
  ggplot(df_f1_chol_refQ3 , aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio") 

#factor 2

df_f2_chol_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(cholf2_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(cholf2_refQ3)[2:24]),
    boxCIHigh = exp(confint(cholf2_refQ3)[2:24, 2])
  )

pf2 <-
  ggplot(df_f2_chol_refQ3 , aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")

#factor 3
df_f3_chol_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(cholf3_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(cholf3_refQ3)[2:24]),
    boxCIHigh = exp(confint(cholf3_refQ3)[2:24, 2])
  )

pf3 <-
  ggplot(df_f3_chol_refQ3 , aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")

#factor 4
df_f4_chol_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(cholf4_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(cholf4_refQ3)[2:24]),
    boxCIHigh = exp(confint(cholf4_refQ3)[2:24, 2])
  )

pf4 <-
  ggplot(df_f4_chol_refQ3 , aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio")

#factor 5
df_f5_chol_refQ3 <-
  data.frame(
    yAxis = length(boxLabels_refQ3):1,
    boxOdds = exp(cholf5_refQ3$coefficients[2:24]),
    boxCILow = exp(confint(cholf5_refQ3)[2:24]),
    boxCIHigh = exp(confint(cholf5_refQ3)[2:24, 2])
  )

pf5 <-
  ggplot(df_f5_chol_refQ3 , aes(x = boxOdds, y = boxLabels_refQ3)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio") 


#all factors with q3 as a reference
boxLabels15_Q3ref = c("Female", "Age", "Ethnicity(Mexican)","Energy" , "< HS/GED" ,"HS/GED", "US < 11 yrs", "US > 10 yrs", "Income < $25k",  "Income $25-75k", "Unmarried",  "Retired", "Unemployed", "Current Drinker", "Former Drinker", "Smoker", "Cholesterol", "Obesity", "Hypertension", "Meats-1st Quintle", "Meats-2nd Quintile", "Meats-4th Quintile", "Meats-5th Quintile", "Fats/Oils-1st Quintle", "Fats/Oils-2nd Quintile", "Fats/Oils-4th Quintile", "Fats/Oils-5th Quintile", "Grains/Legumes-1st Quintle", "Grains/Legumes-2nd Quintile", "Grains/Legumes-4th Quintile", "Grains/Legumes-5th Quintile","Dairy-1st Quintle", "Dairy-2nd Quintile", "Dairy-4th Quintile", "Dairy-5th Quintile", "Fruits/Veggies-1st Quintle", "Fruits/Veggies-2nd Quintile", "Fruits/Veggies-4th Quintile", "Fruits/Veggies-5th Quintile" )


df_chol_f15_Q3ref <-
  data.frame(
    yAxis = length(boxLabels15_Q3ref):1,
    boxOdds = exp(cholf15_Q3ref$coefficients[2:40]),
    boxCILow = exp(confint(cholf15_Q3ref)[2:40, 1]),
    boxCIHigh = exp(confint(cholf15_Q3ref)[2:40, 2])
  )

pf15_Q3ref <-
  ggplot(df_chol_f15_Q3ref, aes(x = boxOdds, y = boxLabels15_Q3ref)) +
  geom_vline(aes(xintercept = 1), size = .25, linetype = "dashed") +
  geom_errorbarh(
    aes(xmax = boxCIHigh, xmin = boxCILow),
    size = .5,
    height =
      .2,
    color = "gray50"
  ) +
  geom_point(size = 1.5, color = "blue") +
  xlim(0, 3.7) +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  ylab("") +
  xlab("Odds ratio") 

ggpubr::ggarrange(pf1, pf2, pf3, pf4, pf5, pf15_Q3ref, labels = c("Meats", "Fats/Oils","Grains/Legumes", "Dairy", "Fruits/Veggies", "All Nutrient Patterns"), ncol = 2, nrow = 3)
```


```{r, echo=F,  fig.width=10, fig.height=10, message=F, warning=F}
#saving results into dataframe to export for comparison
NHANES_Chol_SingleFactors_refQ3 <- cbind(df_f1_chol_refQ3, df_f2_chol_refQ3, df_f3_chol_refQ3, df_f4_chol_refQ3, df_f5_chol_refQ3)

#write.csv(df_chol_f15_Q3ref, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Regression Analysis/NHANES_Chol_AllFactors_ORest_Q3ref_10.23.23.csv")

#write.csv(NHANES_Chol_SingleFactors_refQ3, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Regression Analysis/NHANES_Chol_SingleFactors_ORest_Q3ref_10.23.23.csv")
```

```{r}
#save pvalues in a excel file
ALLadjusted_pvals <- as.data.frame(cbind(diab_adj_pvals_table_all, hyp_adj_pvals_table_all, obs_adj_pvals_table_all, chol_adj_pvals_table_all))

colnames(ALLadjusted_pvals) <- c("Diabetes", "Hypertension", "Obesity", "Cholesterol")

#write.csv(ALLadjusted_pvals, "~/OneDrive - Harvard University/CVD-Nutrient Research/Analysis/Analysis for Both/Regression Analysis/NHANES_pvales_single_full_Q3ref_10.23.23.csv")
```



